### 2015-01-08 material/similiaritylearning.tex

\documentclass{article} % For LaTeX2e
\usepackage{nips11submit_e,times,url,comment,amsmath,amsfonts}
%\documentstyle[nips10submit_09,times,art10]{article} % For LaTeX 2.09

\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\newcommand{\RR}{\mathbb{R}}

\title{Spanning Spaces: Learning Cross-Lingual Similarities}

\begin{comment}
\author{
Jan Rupnik \\
A.I. Laboratory\\
Jo\v zef Stefan Institute\\
Ljubljana, Slovenia\\
\texttt{jan.rupnik@ijs.si} \\
\And
Andrej Muhi\v c\\
A.I. Laboratory\\
Jo\v zef Stefan Institute\\
Ljubljana, Slovenia\\
\texttt{andrej.muhic@ijs.si} \\
\AND
Primo\v z \v Skraba \\
A.I. Laboratory\\
Jo\v zef Stefan Institute\\
Ljubljana, Slovenia\\
\texttt{primoz.skraba@ijs.si} \\
}
\end{comment}
\author{
Jan Rupnik, Andrej Muhi\v c, Primo\v z \v Skraba \\
A.I. Laboratory\\
Jo\v zef Stefan Institute\\
Ljubljana, Slovenia\\
\texttt{(jan.rupnik),(andrej.muhic),(primoz.skraba)@ijs.si} \\
}
% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\X}{\mathbb{X}}
\newcommand{\Y}{\mathbb{Y}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
In analyzing multilingual text corpora, we have the practical
problem of computing similarities between documents in different
languages. Given two documents in different languages, we use
monolingual similarity to an aligned set to compute a similarity
across languages. We derive several algorithms and show their
relationship the choice of similarity function. We also show
experimental results illustrating the approach.
\end{abstract}

\section{Introduction}
Many machine learning tasks such as clustering, classification,
retrieval, etc. rely on computing a metric to compare (data
points). In our motivating example, multilingual corpora, we are
usually faced with an overwhelming imbalance: some languages have
far more documents than others.  Exacerbating the problem is that
the intersection can be small or non-existent. Therefore, we must
use what alignments exist. In this paper, we derive algorithms
based on 2 general approaches. We find that one approach is
approximated by cross-canonical correlation analysis
(CCA)~\cite{Hotelling} and latent semantic indexing (LSI)\cite{lsi} approximates another.

In the next section we derive algorithms based on analyzing the
covariance structure of the aligned data.  We then conclude by
briefly discussing experimental results on the Europarl
dataset~\cite{europarl} and generalizations and future
directions.

\section{Approach}

We begin with a collection of documents in two
languages\footnote{The extension to multiple language is
 discussed in the Discussion section.}. The documents are
represented via bag-of-words as vector spaces $\X$ and
$\Y$. There are several possible choices of monolingual
similarity functions. We consider functions of the form:
$x'C_{\X\X}^i x$ for $i\in \mathbb{Z}$. For $i=0$, we recover the
squared Euclidean distance and for $i=-1$, we get the Mahalanobis
distance.

As input, we take a choice of $i$ (choice of inner product) along
with an alignment of $\X$ and $\Y$. The problem is to find a
mapping which respects both monolingual similarity and measures
cross-lingual similarity.  There are two pragmatic approaches to finding the mapping. In the
first we reduce the problem to monolingual similarity. We
find some mapping $f: \X \rightarrow \Y$. For a pair of elements
$x\in\X$ and $y\in\Y$, we compute their similarity as the
similarity of $f(x)$ and $y$ in $\Y$. The second approach is to
find an aligned joint basis for $\X$ and $\Y$, which can be
interpreted as mapping both elements to a common space and
computing similarity in this common representation.

For the latter approach, we use LSI to find an aligned basis and
for the former we used least-squares regression. We tested these
approaches for 3 choices of similarity: $i=-1,0,1$, corresponding
to Mahalanobis distance, Euclidean distance, and the metric
weighted by the covariance matrix.

The output of all the algorithms is a linear map (a matrix)
denoted $B$, which computes the cross-lingual similarity as $x^T B
y$ for $x\in\X$ and $y \in \Y$. The simplest choice for $B$ is
$C_{\X\Y}$, the cross-covariance matrix.

{\bf LSI:} To compute $B$, we take the first $d$
eigenvalue-eigenvector pairs $\left(\mu_i, [{u_\X^i}^T,
  {u_\Y^i}^T]^T\right)$ of the cross-covariance matrix, we obtain the approximation of the full
covariance matrix:
\begin{equation}
\begin{bmatrix}
  C_{\X \X} & C_{\X \Y} \\
  C_{\Y \X} & C_{\Y \Y}
\end{bmatrix} \approx \sum_{i = 1}^d \mu_i \begin{bmatrix} u_\X^i \\ u_\Y^i \end{bmatrix} \cdot \begin{bmatrix} {u_\X^i}^T & {u_\Y^i}^T \end{bmatrix}, \qquad \qquad B = \sum_{i = 1}^d \mu_i u_\X^i {u_\Y^i}^T\approx C_{\X \Y}
\end{equation}
The definition of $B$ comes from comparing the upper-right
block. This also shows that LSI approximates the simplest choice
of $B$. It is however far more computationally efficient because
it first finds a low-dimensional representation.

{\bf Regression:} To find $B$ using regression, we compute the
optimal regression matrix as $ C_{\Y \X} (C_{\X \X})^{-1}$. This
corresponds to the mapping $f$ mentioned above.  The inner
product on the space $\Y$ induces $B$ for computing the
similarities between elements in $\X$ and $\Y$: for $B$ it holds
that $x^T B y = \langle f (x), y \rangle$. Therefore $B$ depends
on the metric on $\Y$. For the inner product corresponding to
$C_{\Y\Y}$, $B$ can be derived as $C_{\X \X}^{-1}C_{\X
  \Y}C_{\Y \Y}$.

Note that in the case of regression, we have a choice of which
direction we map: $\X\rightarrow\Y$ or $\Y\rightarrow\X$ except
for Mahalanobis metric where the formula for $B$ is
symmetric. Due to space constraints we only note here that CCA approximates the least-squares
regression for this metric just as LSI approximates the
cross-covariance matrix.

\section{Experiments}
To evaluate a cross-similarity matrix performs, we used
monolingual similarity matrices for supervision. All experiments
are based on an aligned bilingual corpus of English and German
documents and all computations were performed by selecting an
aligned subset of $5000$ documents $T_X, T_Y$ for learning $B$
and an aligned set of $500$ documents $R_X, R_Y$
for testing. For the monolingual similarities $\langle \cdot, \cdot
\rangle_X$ and $\langle \cdot, \cdot \rangle_Y$ we compute
monolingual similarity scores: $S_X := \langle R_X, R_X
\rangle_X$ and $S_Y:= \langle R_Y, R_Y \rangle_Y$ and compare the
similarities to $S_{XY} := R_X^T B R_Y$. The measure chosen was
the average correlation coefficient between rows of $S_{XY}$ and
rows of $S_Y$ and the average correlation coefficient between the
columns of $S_{XY}$ and rows of $S_Y$.

We first observe that metrics $C_{X X}^i$ with higher exponent
$i$ are easier to learn (the average correlation coefficients
between the monolingual and cross-lingual similarity profiles are
around $0.9$. We also observe that $i=-1$ is the only approach
that performs relatively well in the inner product space
$C_{XX}^{-1}$ when compared to the other approaches. This is
predicted by the theory since the cross-view similarity matrix
was derived for that objective. We also note that the ad-hoc
approach of $B:= C_{XY}$ and its LSI-based approximation perform
similarly with only $5$ basis vectors used by LSI, which is quite
surprising.

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
&\multicolumn{2}{|c|}{}&\multicolumn{5}{|c|}{Regression} \\
\hline
&       $C_{xy}$	&LSI & 	$i=-1$& $i=0$(xy)& $i=0$(yx) & $i=1$(xy) & $i=1$(yx)\\
\hline
$C_{xx}^{-1}$ &	0.06&	0.05&	0.59&	0.22&	0.21&	0.04&	0.04\\
\hline
$C_{yy}^{-1}$ &	0.06&	0.05&	0.58&	0.22&	0.21&	0.04&	0.04\\\hline
$I_{xx}$     &	0.29&	0.28&	0.70&	0.45&	0.45&	0.27&	0.27\\\hline
$I_{yy}$     &   0.27&	0.26&	0.66&	0.43&	0.42&	0.25&	0.25\\\hline
$C_{xx}$     &	0.94&	0.93&	0.34&	0.87&	0.88&	0.94&	0.94\\\hline
$C_{yy}$     &	0.94&	0.93&	0.33	&0.88&	0.88&	0.95&	0.94\\\hline
\end{tabular}
\end{center}

\bibliographystyle{unsrt} \bibliography{nips}

\end{document}

From now on we assume a fixed $i$ and denote the Hilbert space
corresponding the the spaces and choice of inner product by $X =
\left(\RR^n, \langle\cdot, \cdot\rangle_X\right)$ and $Y =
\left(\RR^m, \langle\cdot, \cdot\rangle_Y\right)$. Let $d_X :(X,
X) \rightarrow \RR$ and $d_Y : (Y,Y) \rightarrow \RR$ be the
induced metrics on $\RR^n$ and $\RR^m$ respectively. Let $X
\oplus Y$ denote the direct sum of Hilbert spaces with the inner
product: $\langle (x_1, y_1), (x_2, y_2) \rangle_{X \oplus Y} =
\langle x_1,x_2 \rangle_X + \langle y_1, y_2 \rangle_Y$. Let $A
\subset X \oplus Y$ be a Riemannian manifold with metric tensor
$g$. Let $\pi_X: X \oplus Y \rightarrow X$ and $\pi_Y: X \oplus Y
\rightarrow Y$ denote projection operators to spaces $X$ and
$Y$. Let $A_X := \pi_X(A)$ and $A_Y := \pi_X(A)$. Let $\pi_{A_X}
: X \rightarrow A_X$ and $\pi_{A_Y} : Y \rightarrow A_Y$ denote
projection operators.

\section{Algorithms}

\section{Partially Aligned Metric Spaces}

We assume our spaces (document corpora divided by language) are
mapped from some underlying, space(language)-independent latent
space denoted $\L$. These mappings differ depending on
language. For simplicity, we concentrate on the case of two
spaces: $\X$ and $\Y$. Therefore, we can write $\X
\overleftarrow{f_x} \L \overrightarrow{f_y} \Y$.  There are
numerous metrics we can choose within a single space: this
depends on what similiarity measure we choose between documents.

The 3 methods we concentrate on are kernel versions of
regression, LSI and CCA~\cite{shawe-taylor04kernel}.  By treating
the aligned documents as mappings of common samples, we can
conclude that there exists a bilinear operator for the metrics
$M_\X \times M_\Y\rightarrow \RR$ and we base this mapping on the
based on covariance matrices. We show how the different methods
correspond with different choices of metrics on the space.

The general statement of our problem is: Find an near-isometric
mapping in between the similiarity spaces for all vectors $x\in
\X$ and $y\in Y$. By working in the covariance space, this
problem reduces to finding transformations between basis vectors
of the two spaces. In all cases, we us the correspondence
information to find this transformation. We limit ourselves to
matrix operator $B(x,y) =  y' B x $

{\bf Regression:} In this case we have the choice of minimizing
$Bx-y$ such that $x$ and $y$ are in correspondence. We can
optimze its adjoint($\Y\rightarrow\X$). In terms of the
covariance matrices we can find $f_{\X\rightarrow\Y}(x) =
C_{\X\X} C_{\Y\Y}^-1 $.  These are obtained by by standard
multivariate regression analysis where the first gives the
optimal regression matrix based on the standard inner
product. Alternatively, if we assume an inner product weighted by
the covariance matrix, the optimal mapping becomes just the
cross-correlation matrix.

\textcolor{red}{DRAWBACKS}

{\bf CCA:}

{\bf LSI:}

In the case of
aligned samples, we have well-defined covariance matrices
$C_{\X\X}$ and $C_{\Y\Y}$.

Notation

Let $L_1 = \RR^{n_1},\ldots, L_m = \RR^{n_m}$ be $m$ vector
spaces. Let $X_1 \in L_1, \ldots, X_m \in L_m$ be continuous
random vectors with defined covariance matrices $C_{X_i X_i} :=
E\big(X_i - E(X_i)(X_i - E(X_i))^T\big)$. In this paper we
consider several possible inner product spaces for $L_1,\ldots,
L_m$ that are induced by powers of covariance matrices. Let
$H_i^p := (L_i, \langle \cdot, \cdot \rangle_{(C_{X_i X_i})^p})$
denote the Hilbert space corresponding to space $L_i$ equipped
with the inner product induced by the matrix $(C_{X_i X_i})^p$.

Assumption

Random vectors $X_i$ are different representations of some
underlying random vector $Z \in R^\ell$: $X_i := f_i(Z), \ldots,
X_m := f_m(Z)$, where each $f_i :\RR^\ell \rightarrow \RR^{n_i}$
is a mapping. For this reason, we will refer to spaces $L_1,
\ldots, L_m$ as different views of some latent objects.

Notation

A dataset $D: = [D_1^T,\ldots, D_m^T]^T $ is an aligned data set
if each $D_i$ can be expressed in the following way: $D_i \in
\RR^{n_i \times k} = f_i([z_1,\ldots, z_k])$ and $[z_1,\ldots,
  z_k] \in \RR^{\ell\times k}$ is a sample of $k$ observation
vectors drawn independently according to distribution of $Z$.

2-view formulation of metric learning

Let $(S_1,S_2)$ denote an aligned sample of $k$ pairs of vectors
($S_1 \in \RR^{n_1 \times k}, S_2 \in \RR^{n_2 \times
  k}$). Assign to $S_1$ a matrix of similarities $M_1^p := S_1^T
C_{X_1 X_1}^p S_1$ and matrix $M_2^p := S_2^T C_{X_2 X_2}^p
S_2$. These two matrices represent similarities computed in
$H_1^p$ and $H_2^p$. Our goal is to find a matrix $B$ that can be
used for computing similarities between vectors in $L_1$ and
vectors in $L_2$ that is in "concordance" with similarities in
$H_1^p$ and $H_2^p$. Define $B_{1,2}^p := S_1^T B S_2$. A good
matrix $B$ for sample $S$ is the one which makes $B_{1,2}^p$
close to $M_1^p$ and $M_2^p$. There are several possible measures
of matrix closeness, a typical choice is the Frobenious
norm. Obviously this is not always achievable, but if the random
variables $X_1$ and $X_2$ (they are involved in drawing samples
$S_1$ and $S_2$) share enough of mutual information, it is
possible, for example in the extreme case where $X_1 = X_2$. This
leads to the following optimization problem. We will refer to the
bilinear mapping $B$ as the cross-view similarity measure.

Optimization problem:

\begin{equation}\label{eq:opt1}
\begin{aligned}
& \underset{B \in \RR^{n_1 \times n_2}}{\text{minimize}}
& & E_{(S_1,S_2)} ||B_{1,2}^p - M_1^p||_F^2 + ||B_{1,2}^p - M_2^p||_F^2 \\
\end{aligned}
\end{equation}

Remark:

$S_1$ and $S_2$ are random matrices (i.i.d. samples of random
variables $X_1$ and $X_2$) and they are sampled in a way that
they form an aligned data set.

Questions:

Can the problem be solved analytically?

Does to choice of the parameter $k$ (sample set size) influence the optimal matrix $B$.

Remark:

In practice we only have a single aligned sample, perhaps this
simplifies the optimization problem (drop the expectation).

Current candidate solutions

For $H_1^0, H_2^0$ we propose: $B := C_{X_1 X_2} C_{X_2 X_2}^{-1}$ or $B := C_{X_1 X_1}^{-1} C_{X_1 X_2}$.
For $H_1^1, H_2^1$ we propose: $B := C_{X_1 X_2}$.
For $H_1^2, H_2^2$ we propose: $B := C_{X_1 X_2} C_{X_2 X_2}$ or $B := C_{X_1 X_1} C_{X_1 X_2}$.

Connection to least squares regression

The candidate solutions can be derived easily if we consider
multivariate regression analysis. The aligned sample $S_1 \in
\RR^{n_1 \times k}$ and $S_2 \in \RR^{n_2 \times k}$ can be
interpreted as a set of input output vector pairs. Using the
method of multivariate least squares regression, an optimal
regression matrix from $L_1$ to $L_2$ can be computed as:
$\widetilde{W} := S_2 S_1^T (S_1 S_1^T)^{-1}$. When $k
\rightarrow \infty$ the expression for $\widetilde{W}$ converges
to $W:= C_{X_2 X_1} (C_{X_1 X_1})^{-1}$. The inner product on the
space $L_2$ induces the bilinear mapping $B$ for computing the
similarities between elements in $L_1$ and $L_2$: the matrix $B$
is the matrix for which it holds $x^T B y = \langle W x, y
\rangle$. For the space $H_2^0$ ($L_2$ equipped with the standard
inner product) we obtain the candidate $B = C_{X_1 X_1}^{-1}
C_{X_1 X_2}$, whereas $H_2^1$ induces $B = C_{X_1 X_1}^{-1}C_{X_1 X_2}C_{X_2 X_2}$. If the
direction of the regression is reversed (consider $S_2$ as the
input vectors and $S_1$ as the output vectors), then the inner
product structure of $L_1$ is the one that induces $B$. This way
we obtain the candidate corresponding to $H_1^0$: $B := C_{X_1
  X_2} C_{X_2 X_2}^{-1} $. The possible problem with the
regression approach to similarity induction is that only one
inner product structure is involved in computing $B$, whereas the
search for $B$ in \ref{eq:opt1} considers both inner product
spaces.

Three views

Assume that we have three views $X_1,X_2$ and $X_3$. If $C_{X_1 X_3}$ is not available (too few data points to estimate) we propose:
$B := C_{X_1 X_2} C_{X_2 X_3}$ which fits inner product spaces $H_1^2$ and $H_3^2$.

Long chains are inappropriate practically because they correspond to inner product spaces induced by variance matrices exponentiated to high powers.
Rank of such matrices drops quickly (it approaches the dimension of the dominant eigen-space of the variance matrix) with the exponent and they typically become unsuitable for practical applications.

CCA, SVD and covariances

CCA and SVD represent two forms of analysis of data based on the covariance structures. They can both be used to extract basis vectors for each view that can be used for computing the bilinear mapping that enables computations of cross-view similarity mappings.

Assume that $m=2$.

Canonical correlation analysis can be formulated as following problem:

\begin{equation*}
\begin{aligned}
& \underset{w_1 \in L_1, w_2 \in L_2}{\text{maximize}}
& &  \frac{w_1' C_{X_1 X_2} w_2}{\sqrt{w_1' C_{X_1 X_1} w_1} \sqrt{w_2' C_{X_2 X_2} w_2}}.
\end{aligned}
\end{equation*}

This can be reformulated as:
\begin{equation}\label{eq:cca}
\begin{aligned}
& \underset{w_1 \in L_1, w_2 \in L_2}{\text{maximize}}
& &  w_1' C_{X_1 X_2} w_2\\
& \text{subject to}
& &w_1' C_{X_1 X_1} w_1 = 1
& &w_2' C_{X_2 X_2} w_2 = 1.
\end{aligned}
\end{equation}

Using lagrangian multipliers technique yields a generalized eigenvalue problem:
$$
\begin{bmatrix}
  0 & C_{X_1 X_2} \\
  C_{X_2 X_1} & 0 \\
\end{bmatrix} \cdot
\begin{bmatrix}
  w_1 \\
  w_2 \\
\end{bmatrix} =
\lambda
\begin{bmatrix}
  C_{X_1 X_1} &0  \\
  0 & C_{X_2 X_2} \\
\end{bmatrix} \cdot
\begin{bmatrix}
  w_1 \\
  w_2 \\
\end{bmatrix}
$$

By inverting the matrices $C_{X_1 X_1}$ and $C_{X_2 X_2}$ we arrive to a standard eigenvalue problem:

$$
\begin{bmatrix}\label{eq:cca_ev}
  0 & C_{X_1 X_1}^{-1} C_{X_1 X_2} \\
  C_{X_2 X_2}^{-1} C_{X_2 X_1} & 0 \\
\end{bmatrix} \cdot
\begin{bmatrix}
  w_1 \\
  w_2 \\
\end{bmatrix} =
\lambda
\begin{bmatrix}
  w_1 \\
  w_2 \\
\end{bmatrix}
$$

We then compute the first $d$ largest eigenvectors and their eigenvalues $(\lambda_i, w_1^i, w_2^i),~i=1,\ldots,d$, where $w_1^i$ and $w_2^i$ are scaled so that constraints in \ref{eq:cca} are satisfied: ${w_1^i}^T C_{X_1 X_1} w_1^i = 1$ and ${w_2^i}^T C_{X_2 X_2} w_2^i = 1$.

Combining the constraints with $C_{X_1 X_1}^{-1} C_{X_1 X_2} w_2^i = \lambda_i w_1^i$ from \ref{eq:cca_ev} we observe a close connection between multivariate least squares regression and CCA canonical vectors:
$$C_{X_1 X_1}^{-1} C_{X_1 Y_1} = \left(\sum_{i=1}^d \lambda_i w_1^i {w_2^i}^T\right)C_{X_2 X_2}$$
This means that the basis vectors can be used to form a cross-view similarity mapping:
$$B = \sum_{i=1}^d \lambda_i w_1^i {w_2^i}^T= C_{X_1 X_1}^{-1} C_{X_1 Y_1} C_{X_2 X_2}^{-1}.$$

We have motivated the use of bilinear mapping that seems less biased than the LSR regression variants where the direction of the regression matters.

SVD

Singular value decomposition is a popular approach in dimensionality reduction. It can be readily applied to the cross-view similarity learning in the following way. An aligned sample $\left(S_1, S_2\right)$ of $k$ pairs of observation vectors is analyzed in the following way.

Stack the observations into a matrix of dimensions $k$-by$n_1 +n_2$:
$$
S = \begin{bmatrix} S_1 \\ S_2 \end{bmatrix}
$$
Compute the singular value decomposition of $S = U \Sigma V^T$ and retain the first $d$ columns of $U$ that correspond to the $d$ largest singular values in $\Sigma$. The vectors $U$ can be computed by solving following eigenvalue problem:
$$
\begin{bmatrix}
  S_1 S_1^T & S_1 S_2^T \\
  S_2^T S_1 & S_2 S_2^T \\
\end{bmatrix} \cdot
\begin{bmatrix}
  u_1 \\
  u_2 \\
\end{bmatrix} =
\mu
\begin{bmatrix}
  u_1 \\
  u_2 \\
\end{bmatrix}.
$$
Here we decomposed the eigenvectors to fit the block structure imposed by $S_1$ and $S_2$. Dividing both sides of the equation by the sample size and sending the sample size to infinity corresponds to the optimization on the covariance matrices:
$$
\begin{bmatrix}
  C_{X_1 X_1} & C_{X_1 X_2} \\
  C_{X_2 X_1} & C_{X_2 X_2} \\
\end{bmatrix} \cdot
\begin{bmatrix}
  u_1 \\
  u_2 \\
\end{bmatrix} =
\mu
\begin{bmatrix}
  u_1 \\
  u_2 \\
\end{bmatrix}.
$$
Taking the first $d$ eigenvalue-eigenvector pairs $\left(\mu_i, [{u_1^i}^T, {u_2^i}^T]^T\right)$  we obtain the approximation of the full covariance matrix:
$$
\begin{bmatrix}
  C_{X_1 X_1} & C_{X_1 X_2} \\
  C_{X_2 X_1} & C_{X_2 X_2} \\
\end{bmatrix} \approx \sum_{i = 1}^d \mu_i \begin{bmatrix} u_1^i \\ u_2^i \end{bmatrix} \cdot \begin{bmatrix} {u_1^i}^T & {u_2^i}^T \end{bmatrix}
$$
Comparing the upper-right block shows:
$$C_{X_1 X_2} \approx \sum_{i = 1}^d \mu_i u_1^i {u_2^i}^T.$$
We now see that the SVD based approach yields a cross-view similarity mapping $B = \sum_{i = 1}^d \mu_i u_1^i {u_2^i}^T$ that is an approximation of the regression induced bilinear similarity mapping on inner product space $H_1^1$ (or equivalently $H_2^1$).

Kernel CCA, kernel SVD

In order to apply CCA and SVD to Hilbert spaces other than $H_i^0$ we used their kernel counterparts. The modified methods are described in \cite{shawe-taylor04kernel}.

Corrections: leveraging unaligned data

All involved methods were operating on covariance estimates based on aligned samples. In practice we often encounter the situation where some data is aligned but each view contains additional unaligned data. Such data can be used to obtain better estimates of variance matrices $C_{X_i X_i}$. Let $m=2$. Let $\widetilde{S}_1 \in \RR^{n_1 \times s_1}$ represent a sample of the random variable $X_1$ of size $s_1$, $\widetilde{S}_2 \in \RR^{n_2 \times s_2}$ a sample of random variable $X_2$ of size $s_2$ and $\left(\widehat{S}_1 \subset \widetilde{S}_1, \widehat{S}_2 \subset \widetilde{S}_2\right)$ an aligned sample of size $k$. Assume that the random vectors are centered. We can obtain the following empirical estimates: $\widetilde{C}_{X_1 X_1} := \frac{1}{s_1 -1} \widetilde{S}_1 \widetilde{S}_1^T$, $\widetilde{C}_{X_2 X_2} := \frac{1}{s_2 -1} \widetilde{S}_2 \widetilde{S}_2^T$, $\widehat{C}_{X_1 X_1} := \frac{1}{k -1} \widehat{S}_1 \widehat{S}_1^T$, $\widehat{C}_{X_2 X_2} := \frac{1}{k -1} \widehat{S}_2 \widehat{S}_2^T$ and $\widehat{C}_{X_1 X_2} := \frac{1}{k -1} \widehat{S}_1 \widehat{S}_2^T$. All of the mentioned approaches to finding the bilinear form are based on computations involving the empirical estimates. Computations based on replacing $\widehat{C}_{X_1 X_1}$ with $\widetilde{C}_{X_1 X_1}$ and  replacing$\widehat{C}_{X_2 X_2}$ with $\widetilde{C}_{X_2 X_2}$ typically perform worse than results based on $\widehat{C}_{X_1 X_1}$, $\widehat{C}_{X_2 X_2}$ and $\widehat{C}_{X_1 X_2}$ alone.

The approach we propose is to compute the change of basis that transforms $\widehat{C}_{X_1 X_1}$ to $\widetilde{C}_{X_1 X_1}$, $\widehat{C}_{X_2 X_2}$ to $\widetilde{C}_{X_2 X_2}$ and apply it to $\widehat{C}_{X_1 X_2}$. To achieve this we first compute Cholesky decompositions of $\widetilde{C}_{X_1 X_1} = \widetilde{C}_1^T \widetilde{C}_1$, $\widetilde{C}_{X_2 X_2} = \widetilde{C}_2^T \widetilde{C}_2$, $\widehat{C}_{X_1 X_1} = \widehat{C}_1^T \widehat{C}_1$, $\widehat{C}_{X_2 X_2} = \widehat{C}_2^T \widehat{C}_2$. The matrices that perform the required change of basis are: $\widetilde{C}_1 \widehat{C}_1^{-1}$ and $\widetilde{C}_2 \widehat{C}_2^{-1}$. The new input to our methods, the variance and covariance matrices, are: $\widetilde{C}_{X_1 X_1}$, $\widetilde{C}_{X_2 X_2}$ and $\widetilde{C}_1^T \widehat{C}_1^{-T}  \widehat{C}_{X_1 X_2} \widehat{C}_2^{-1} \widetilde{C}_2$.

Alternative problem definition: differential geometric approach

Let $X = \left(\RR^n, \langle\cdot, \cdot\rangle_X\right)$ and $Y
= \left(\RR^m, \langle\cdot, \cdot\rangle_Y\right)$ be two
Hilbert spaces referred to as views and let $d_X :(X, X)
\rightarrow \RR$ and $d_Y : (Y,Y) \rightarrow \RR$ be the induced
metrics on $\RR^n$ and $\RR^m$ respectively. Let $X \oplus Y$
denote the direct sum of Hilbert spaces where the inner product
is computed as $\langle (x_1, y_1), (x_2, y_2) \rangle_{X \oplus
  Y} = \langle x_1,x_2 \rangle_X + \langle y_1, y_2
\rangle_Y$. Let $A \subset X \oplus Y$ be a Riemannian manifold
with metric tensor $g$. Let $\pi_X: X \oplus Y \rightarrow X$ and
$\pi_Y: X \oplus Y \rightarrow Y$ denote projection operators to
spaces $X$ and $Y$. Let $A_X := \pi_X(A)$ and $A_Y :=
\pi_X(A)$. Let $\pi_{A_X} : X \rightarrow A_X$ and $\pi_{A_Y} : Y
\rightarrow A_Y$ denote projection operators.

Idealized assumption: existence of diffeomorphisms $f_X : A_X \rightarrow A$ and $f_Y : A_Y \rightarrow A$.

Define the cross-view measure of distance as the mapping $B_A : (A_X, A_Y) \rightarrow \RR$,
$$B(x,y) := \int_0^1 \sqrt{ g_{\gamma(t)}\left(\dot{\gamma}(t), \dot{\gamma}(t)\right)}, $$ where $x \in A_X$, $y \in A_Y$, $\gamma$ is a geodesic curve lying in $A$ and $\gamma(0) = \left( x, f_X(x)\right)$ and $\gamma(1) = \left( f_Y(y), y\right)$.

The interpretation of the space $A$ is that it encodes the mutual information between views. An aligned sample $S = (S_X, S_Y) \subset X \oplus Y$ represents a collection of points lying in the space $A$.

 Extending the cross-view measure of distance from $A_X \oplus A_Y$ to full $X \oplus Y$ is straightforward: for $x \in X$ and $y \in Y$ define $B(x,y):= B_A\left(\pi_{A_X}(x), \pi_{A_Y}(y)\right)$.

If $A$ is a linear space

\section{Algorithms}

\noindent {\bf Baseline:}

\noindent {\bf Regression:}

\noindent {\bf CCA with correction:}

\noindent {\bf Global Methods:}

\label{sec:alg}
\section{Experimentation}

\section{Discussion}

\bibliographystyle{unsrt} \bibliography{nips}

\end{document}