### 2015-07-20 rupnik2015.tex
updated the pipeline section. added a better image and removed two existing ones. added 3 references

\documentclass[twoside,11pt]{article}
\usepackage{jair, theapa, rawfonts}
\usepackage{pgfplots}
%\usepackage{graphicx}

\usepackage{color}
\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
\newcommand{\oldText}[1]{\textcolor{red}{#1:}\color{blue}}

\usepackage[implicit=false]{hyperref}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{comment}

\usepackage[]{algorithm2e}

\newcommand{\RR}{\mathbb{R}}
\newcommand{\X}{\mathbb{X}}
\newcommand{\Y}{\mathbb{Y}}

\newcommand{\algInput}{\textbf{input:}}

\newcommand{\primoz}[1]{\textcolor{red}{#1}}

\jairheading{x}{xxxx}{x-xx}{x/xx}{x/xx}
\ShortHeadings{Cross-Lingual Document Similarity and Event Tracking}
{Rupnik, Muhi\v{c}, Leban, Fortuna, \v{S}kraba \& Grobelnik}
\firstpageno{1}

\begin{document}

\title{News Across Languages -
Cross-Lingual Document Similarity and Event Tracking}

\author{\name Jan Rupnik \email jan.rupnik@ijs.si \\
       \name Andrej Muhi\v{c} \email andrej.muhic@ijs.si \\
       \name Gregor Leban \email gregor.leban@ijs.si \\
       \name Primo\v{z} \v{S}kraba \email primoz.skraba@ijs.si \\
       \name Bla\v{z} Fortuna \email blaz.fortuna@ijs.si \\
       \name Marko Grobelnik \email marko.grobelnik@ijs.si \\
       \addr Artificial Intelligence Laboratory, Jo\v{z}ef Stefan Institute,\\
       Jamova cesta 39, 1000 Ljubljana, Slovenia}

\maketitle

\begin{abstract}
Today we follow news which is distributed globally. Significant events are reported by different sources and in different languages. In this work, we address the problem of tracking and events in a large multilingual stream. We consider a particular aspect of this problem, namely how to link collections of articles in different languages which refer to the same event. Given a multi-lingual stream and clusters of articles from each language, we first propose a method for cross-lingual document similarity based on Wikipedia, which enables us to compute the similarity of any two articles regardless of language. The proposed method can scale to 100 languages and can match articles from languages with little or no direct overlap in the training data. Using this similarity, we then propose an approach to link clusters of articles across languages which represent the same event. We provide an extensive evaluation of the system as a whole, as well as an evaluation of the quality and robustness of the similarity measure and the linking algorithm.
\begin{comment}
Main points:
\begin{itemize}
\item Problem of tracking events in a large multilingual stream - we reduce it to event linking
\item Two contributions. We propose a method that enables cross-lingaul document similarity an approach to linking multilingual data and an approach that uses is to link events.
\item Evaluation of cross-lingual approaches on Wikipedia and the event linking task.
\end{itemize}
\end{comment}
\end{abstract}
%\section{Overview}
%Structure
%\begin{itemize}
%\item Introduction
%\item Cross lingual document similarity (model)
%\item  Event Detection
%\item  Combing the approaches
%\item Experiments
%\item Conclusions
%\end{itemize}
%The story:
%\begin{itemize}
%\item Description of the problem - we would liek to understand events ans sequences of events, but our observables (e.g. news stories) occur in a variety of languages
%\item We present an approach to tackle this problem using a combination of techniques....
%\item Ultimately we present a tool showing the feasibility of our approach
%%\end{itemize}
%
%\section{Overview}
%Structure
%\begin{itemize}
%\item Introduction
%\item System overview
%\item Cross lingual document similarity (algorithms)
%\item Event linking
%\item Experiments
%\item Conclusions
%\end{itemize}
%The story:
%\begin{itemize}
%\item Description of the problem - we would like to understand global news as sequences of events, but our observables (e.g. news stories) occur in a variety of languages
%\item We present an approach to tackle this problem using a combination of techniques....
%\item Ultimately we present a tool showing the feasibility of our approach
%\end{itemize}
%

\section{Introduction}

Content on the internet is becoming increasingly multi-lingual. A prime example is Wikipedia - in 2001 the majority of pages were written in English, while in 2015, the percentage of English articles has dropped to 14\%. Machine translation remains relatively rudimentary - allowing people to understand simple phrases on web pages, but remain inadequate for more advanced understanding of text. At the same time, online news has begun to dominate reporting of current events. In this paper we consider the intersection of these developments: how to track events which are reported about in multiple languages.

The term event is vague and ambiguous, but for the practical purposes, we define it as ``any significant happening that is being reported about in the media.'' Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015 (Figure \ref{fig:event2}). Events such as these are covered by many articles and the question is how to find all the articles in different languages that are describing a single event.

\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{events2.png}
\caption{\label{fig:event2} Events are represented by collections of articles about an event, in this case the Malaysian airliner which was shot down over the Ukraine.}
\end{figure}

As input, we consider a stream of articles in different languages and a list of events. Our goal is to assign articles to their corresponding events. A priori, we do not know the coverage of the articles, that is not all the events may be covered and we do not know that all the articles necessarily fit into one of the events. The task is divided into two parts: detecting events within each language and then linking events across languages. In this paper we address the second step.

As input we consider a high volume of articles in different languages. By using a language detector the stream is split into separate monolingual streams. Within each monolingual stream, an online clustering approach is employed, where tracked clusters correspond to our definition of events - this is based on the Event Registry system \shortcite{Leban2014W,Leban2014I}. Our main goal in this paper is to connect such clusters (representations of events) across languages, that is detect that a set of articles in language $A$ reports on the same event as a set of articles in language $B$.

Our approach to linking clusters across languages combines two ingredients: a cross-lingual document similarity measure, which can be interpreted as a language independent topic model, and semantic annotation of documents, which enables an alternative way to comparing documents.

The first approach represents a continuation of previous work \shortcite{nips,sikdd,nips2,iti} where we explored representations of documents which were valid over multiple languages.  The representations could be interpreted as multi-lingual topics, which were then used as proxies to compute cross-lingual similarities between documents. To learn the representations, we use Wikipedia as a training corpus. Significantly, we do not only consider the major or \emph{hub} languages such as English, German, French, etc. which have significant overlap in article coverage, but also smaller languages (in terms of number of Wikipedia articles) such as Slovenian and Hindi, which may have a negligible overlap in article coverage. We can then define a similarity between any two articles regardless of language,  which allows us to cluster the articles according to topic. The underlying assumption being that articles describing the same event are similar  and will therefore be put into the same cluster.

Based on the similarity function we propose a novel algorithm for linking events/clusters across languages. The approach is based on learning a classification model from labelled data based on several sets of features. In addition to these features, cross-lingual similarity is also used to quickly identify a small list of potential linking candidates for each cluster.
This greatly increases the scalability of the system.

%As input we consider a stream of articles in different languages and a list of events. Our goal is to assign articles to their corresponding events. A priori we do not know the coverage of the articles, that is not all the events may be covered and we do not know that all the articles necessarily fit into one of the events.  To achieve our goal, we divide the task into two parts: clustering the articles according to topic (regardless of language) and matching the clusters to events.
%
%To cluster the articles, rather than rely on machine translations, we adopt well-established monolingual text mining tools to the multilingual setting. This represents the continuation of previous work \shortcite{nips,sikdd,nips2,iti} where we explored representations of documents which were valid over multiple languages.  The representations could be interpreted as multi-lingual topics, which were then used as proxies to compute cross-lingual similarities between documents. To learn the representations, we use Wikipedia as a training corpus. Significantly, we do not only consider the major or \emph{hub} languages such as English, German, French, etc. which have significant overlap in article coverage, but also smaller languages (in terms of number of Wikipedia articles) such as Slovenian and Hindi, which may have a negligible overlap in article coverage. We can then define a similarity between any two articles regardless of language,  which allows us to cluster the articles according to topic. The underlying assumption being that articles describing the same event are similar  and will therefore be put into the same cluster.
%
%The second step is to match our computed clusters to events. Since our similarity measure and clustering
%may not be perfect, we need to find the ``best" correspondence between clusters and events. In this paper, we test several different algorithms for both steps for performance in an integrated system. In particular, we test our approach within an online system for detection of world events, called Event Registry \shortcite{Leban2014W,Leban2014I}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles.

%
%Event Registry works by collecting and analyzing news articles that are published by news outlets in different languages all over the world. Collected articles are first semantically annotated by identifying in them mentions of relevant concepts -- either entities or important keywords. The disambiguation and entity linking of the concepts is done using Wikipedia as the main knowledge base.
%   The key step in Event Registry is then to apply an online clustering algorithm on the articles in order to identify groups of articles that are discussing the same event.
%
%   Within this system, we test our multilingual similarity measures as well as our algorithms for matching clusters across languages.  The result is that we can test total performance in a multilingual, real-world setting.

The paper is organized as follows: we first provide an overview of the system as a whole (Section ~\ref{sec:pipeline}), followed by details on each of the two steps described above. We first introduce the problem of cross-lingual similarity computation (Section~\ref{sec:crosslingual}), which is followed by several approaches for measuring cross-lingual similarity (Section~\ref{sec:models}) and then the algorithms for matching clusters in different languages (Section~\ref{sec:linking}).  To improve readability, we describe relevant related work for each module in its corresponding Section. Finally, we present and interpret the experimental results, give conclusions and describe promising future directions.

\section{Pipeline}\label{sec:pipeline}

We base our techniques of cross-lingual event linking on an online system for detection of world events, called Event Registry \shortcite{Leban2014W,Leban2014I}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles that are collected from numerous news outlets all over the world. The important components in the pipeline of the Event Registry are shown in Figure~\ref{fig:erpipeline} and will now be briefly described.

\begin{figure}[tbp]
\centering
\includegraphics[width=\textwidth]{pipeline}
\caption{\label{fig:erpipeline}  The Event Registry pipeline. After new articles are collected, they are first analyzed individually (Article-level processing). In the next step, groups of articles about the same event are identified and relevant information about the event is extracted (Event construction phase). Although the pipeline contains several components, we will in this paper focus only on the two highlighted in the image.}
\end{figure}

The collection of the news articles is performed using the Newsfeed service~\cite{Trampus2012}. The service monitors RSS feeds of around 100.000 mainstream news outlets available globally. Whenever a new article is detected in the RSS feed, the service downloads all available information about the article and sends the article through the pipeline. Newsfeed downloads daily on average around 200.000 news articles in various languages, where English, Spanish and German are the most common.

Collected articles are first semantically annotated by identifying in them mentions of relevant concepts -- either entities or important keywords. The disambiguation and entity linking of the concepts is done using Wikipedia as the main knowledge base. The algorithm for semantic annotation is based on~\cite{zhang2014saaacamactat}. As a part of the semantic annotation we also analyze the dateline of the article to identify the location of the described event as well as extract dates mentioned in the article using a set of regular expressions. Since articles are frequently revised we also detect if the collected article is just a revision of a previous one so that we can use this information in next phases of the pipeline. The last important processing step on the document level is to efficiently identify which articles in other available languages are most similar to this article. The methodology for this task is one of the main contributions of this paper and is described in details in Section~\ref{sec:crosslingual}).

As a next step, an online clustering algorithm is applied on the articles in order to identify groups of articles that are discussing the same event. For each new article, the clustering algorithm determines if the article should be assigned to some existing cluster or into a new cluster. The underlying assumption is that articles that are describing the same event are similar enough and will therefore be put into the same cluster. For clustering, each new article is first tokenized, stop words are removed and words are stemmed.
The remaining tokens are represented in a vector-space model and normalized using TF-IDF (see Section~\ref{sec:tfidf} for the definition). The cosine similarity is used to find the most similar existing cluster, by comparing the document's vector to the centroid vector of each cluster. A user-defined threshold is used to determine if the article is not similar enough to any existing clusters (0.4 was used in our experiments).  If the highest similarity is above the threshold, the article is assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article. Since articles about an event are commonly written only for a short period of time, we remove clusters once the oldest article in the cluster becomes more than 4 days old. This housekeeping mechanism prevents the clustering from becoming slow and also ensures that articles are not assigned to obsolete clusters. Details of the clustering approach are described in~\cite{brank2014}.

Once the number of articles in a cluster reaches a threshold (which is a language dependent parameter), we assume that the articles in the cluster are describing an event. At that point, a new event with a unique ID is created in Event Registry, and the cluster with the articles is assigned to it. By analyzing the articles we extract the main information about the event, such as the event location, date, most relevant entities and keywords, etc.

Since articles in a cluster are in a single language, we also want to identify any other existing clusters that are reporting about the same event in another language and join these clusters into the same event. This task is performed using a classification approach that is the second major contribution of this paper. It is described in details in Section~\ref{sec:linking}.

When a cluster is identified and information about the event is extracted, all available data is stored in a custom-built database system. The data is then accessible through the API or a web interface\footnote{http://eventregistry.org/}, which provide numerous search and visualization options.

\primoz{we may want to move this out to the pipeline section}
\subsection{System Requirements}

Our goal is to build a system that monitors global media and analyzes how events are being reported on. Our approach consists of two steps: tracking events separately in each language (based on language detection and an online clustering approach) and then connecting them. The pipeline must be able to process millions of articles per day and perform billions of similarity computations each day - implying that similarity computation should be highly scalable.  %The system should support as many languages as possible.
 We focus on implementations that run on a single shared memory machine, as opposed to clusters of machines. This  simplifies implementation and system maintenance.
To summarize, the following properties are desirable:
\begin{itemize}
\item \textbf{Training} - The training (building cross-lingual models) should scale to many languages and should be robust to the quality of training resources. The system should be able to take advantage of comparable corpus (as opposed to parallel translation based corpora), with missing data. Supporting a new language should be straightforward.
\item \textbf{Operation efficiency} - The similarity computation should be fast - the system must be able to handle billions of similarity computations per day.
\item \textbf{Operation cost} - The system should run on a strong shared machine server and not rely on paid services.
\item \textbf{Implementation} - The system is straightforward to implement, with few parameters to tune.
\end{itemize}
We believe that a cross-lingual similarity component that meets such requirements is very desirable in a commercial setting, where several different costs have to be taken into consideration.

\subsection{Related work}

%In this section we will relate the method we use to other alternatives in the literature. We will first state some requirements for our system, which motivate our choice of focusing on a particular subset of approaches in the rest of the paper.

The goal is to build a system that monitors global media and analyzes how events are being reported on. In our approach this boils down to two steps: tracking events separately in each language (based on language detection and an online clustering approach) and then connecting them. The pipeline needs to process millions of articles per day and perform billions of similarity computations each day. The system should support as many languages as possible. We focus on implementations that run on a single shared memory machine, as opposed to clusters of machines. This severely simplifies the software implementation and system maintenance.
To summarize, the following properties are desireable:
\begin{itemize}
\item \textbf{Training} - The training (building corss-lingual models) should scale to many languages and should be robust to the quality of training resources. The system should be able to take advantage of comparable corpus (as opposed to parallel translation based corpora), with missing data. \primoz{We could say something more subject to moving this section}%Supporting a new language should be straightforward.
\item \textbf{Efficiency} - The similarity computation should be fast - the system must be able to handle billions of similarity computations per day.
\item \textbf{Cost} - The system should run on a strong shared machine server and not rely on paid services.
\item \textbf{Implementation} - The system is simple to implement, with few parameters to tune.
\end{itemize}
We believe that a cross-lingual similarity component that meets such requirements is very desirable in a commercial setting, where several different costs have to be taken into consideration.

There are four main families of approaches to cross-lingual similarity.
\begin{itemize}
\item \textbf{Translation} - The most obvious way to compare documents written in different languages is to use machine translation and perform monolingual similarity. One can use free tools such as Moses \shortcite{moses} or translation services, such as Google Translate\footnote{https://translate.google.com/}. There are two issues with such approaches: they solve a harder problem than needs to be solved and they are less robust to training resource quality - large sets of translated sentences are typically needed. Training Moses for languages with scarce linguistic resources is thus problematic. The issue with using online services such as Google Translate is that the APIs are limited and not free. The operation efficiency and cost requirements make translation based approaches less suited for our system.
\item \textbf{Probabilistic topic models} -  There exist many variants to modelling documents in a language independent way by using probabilistic graphical models. The models include:  Joint Probabilistic Latent Semantic Analysis (JPLSA) \shortcite{platt2010translingual}, Coupled Probabilistic LSA (CPLSA) \cite{platt2010translingual}, Probabilistic Cross-Lingual LSA (PCLLSA)\cite{PCL_LSA} and Polylingual Topic Models (PLTM) \shortcite{polyLDA} which is a Bayesian version of PCLLSA. The methods (except for CPLSA) describe the multilingual document collections as samples from generative probabilistic models, with variations on the assumptions on the model structure. The topics represent latent variables that are used to generate observed variables (words), a process specific to each language. The parameter estimation is posed as an inference problem which is typically intractable and one usually solves it using approximate techniques. Most variants of solutions are based on Gibbs sampling or Variational Inference, which are nontrivial to implement and may require an experienced practitioner to apply. Furthermore, representing a new document as a mixture of topics is another potentially hard inference problem which must be solved.
\item \textbf{Classification problem related} - In approaches such as \cite{cotrain} and \shortcite{svm2k} the authors find linear embeddings of the data that support good classification results, based on a given classification problem training data. We do not consider this family of approaches, since for general language pairs classifier training sets may not be available.
\item \textbf{Matrix factorization} - Non-negative matrix factorization based \cite{nonnegfactor_lsi}, Cross-Lingual Latent Semantic Indexing CLLSI\shortcite{cl_lsi}, Canonical Correlation Analysis (CCA) \cite{Hotelling}, Oriented Principal Component Analysis \cite{platt2010translingual}. The quadratic time and space dependency of the OPCA method makes it impractical for large scale purposes. In addition, OPCA forces the vocabulary sizes for all languages to be the same, which is less intuitive. For our setting, the method in \cite{nonnegfactor_lsi} has a prohibitively high computational cost when building models (it uses dense matrices of dimensions that is the product of training set size and vocabulary size). Our proposed approach combines CCA and CLLSI, which will be presented in more detail.

\end{itemize}

Based on our requirements in Section~\ref{} \primoz{System requirements}, we chose to focus on methods based on vector space models and linear embeddings. We propose a method that is more efficient than popular alternatives (a clustering based approach and latent semantic indexing), but is still simple to optimize and use.

\subsection{Related work}

Although there are a number of services that aggregate news by identifying clusters of similar articles, there are almost no services that provide linking of clusters over different languages. Google News as well as Yahoo! News are able to identify clusters of articles about same event, but they offer no linking of clusters across languages. The only service that we found, which provides cross-lingual cluster linking, is the European Media Monitor (EMM)~\cite{Steinberger2008}. EMM clusters articles in 60 languages and then tries to determine which clusters of articles in different languages describe the same event. To achieve cluster linking, EMM uses three different language independent vector representations for each cluster. First vector contains the weighted list of references to countries mentioned in the articles, while the second vector contains the weighted list of mentioned people and organizations. The last vector contains the weighted list of Eurovoc subject domain descriptors. These descriptors are topics, such as \emph{air transport}, \emph{EC agreement}, \emph{competition} and \emph{pollution control} into which articles are automatically categorized~\shortcite{Pouliquen2006}. Similarity between clusters is then computed using a linear combination of the cosine similarities computed on the lists. If the similarity is above the threshold, the clusters are linked. Compared to EMM, our approach uses document similarities to obtain a small set of potentially equivalent clusters. Additionally, we do not decide if two clusters are equivalent based on a hand-set threshold on a similarity value -- instead we use a classification model that uses a larger set of features related to the tested pair of clusters.

A system that is significantly different, but worth mentioning is the GDELT project~\cite{Leetaru2013Gdelt}. In GDELT, events are also extracted from articles, but in their case, an event is specified in a form of a triple containing two actors and a relation. The project contains an extensive vocabulary of possible relations, mostly related to political events. In order to identify events, GDELT collects articles in more than 65 languages and uses machine translation to translate them to English. All information extraction is then done on the translated article.

\section{Cross-lingual Document Similarity}
\label{sec:crosslingual}
Document similarity is an important component in techniques from text mining and natural language processing. Many techniques use the similarity as a black box, i.e., a kernel in Support Vector Machines. Comparison of documents (or other types of text snippets) in a monolingual setting is a well studied problem in the field of information retrieval \cite{Salton88term-weightingapproaches}. We first recount some system requirements, then review  related approaches to cross-lingual similarity models and then formally introduce the problem as well as our approach.

\subsection{Problem definition}\label{sec:tfidf}
We will first describe how documents are represented as vectors and how to compare douments in a mono-lingual setting. We will then define a way to measure cross-lingual similarity that is natural for the models we consider.

\textbf{Document representation.}
The standard vector space model \cite{Salton88term-weightingapproaches} represents documents as vectors, where each term corresponds to word or phrase in a fixed vocabulary. Formally, document $d$ is represented by a vector $x \in \RR^n$, where $n$ corresponds to the size of the vocabulary, and vector elements $x_k$ correspond to the number of times term $k$ occurred in the document, also called \emph{term frequency} or $TF_k(d)$.

We also used a term re-weighting scheme that adjusts for the fact that some words occur more frequently in general. A term weight should correspond to the importance of the term for the given corpus. The common weighting scheme is called \emph{Term Frequency Inverse Document Frequency} ($TFIDF$) weighting. An \emph{Inverse Document Frequency} ($IDF$) weight for the dictionary term $k$ is defined as $\log\left( \frac{N}{DF_k} \right)$, where $DF_k$ is the number of documents in the corpus which contain term $k$.  Therefore we can definte a document's $TFIDF$ as
$$ x_{ij}  := \frac{\mbox{term frequency in document } i}{\mbox{inverse document frequency of term } j}.$$
The $TFIDF$ weighted vector space model document representation corresponds to a map $\phi : \text{text} \rightarrow \RR^n$ defined by:
$$\phi(d)_k = {TF}_k(d) \log\left( \frac{N}{{DF}_k}\right).$$

\textbf {Mono-lingual similarity.}
A common way of computing similarity between documents is \emph{cosine similarity},
$$sim(d_1, d_2) = \frac{\langle \phi(d_1), \phi(d_2)\rangle}{\|\phi(d_1)\| \|\phi(d_2)\|},$$
where $\langle \cdot,\cdot \rangle$ and $\|\cdot\|$ are standard inner product and Euclidean norm. Cosine similarity, and other related approaches, assumes that the similarity is reflected in the overlap of words. This metric is only meaningful when the documents $d_1$ and $d_2$ are defined on the same space (e.g. have a comparable vocabulary or are written in the same language).

\textbf {Cross-lingual similarity.}
Processing a multilingual dataset results in several vector spaces with varying dimensionality, one for each language. The dimensionality of the vector space corresponding to the $i$-th language is denoted by $n_i$ and and the vector space model mapping is denoted by $\phi_i : \text{text} \rightarrow \RR^{n_i}$.
The similarity between documents in language $i$ and language $j$ is defined as a bilinear operator represented as a matrix $S_{i,j} \in \RR^{n_i \times n_j}$:
$$sim_{i,j}(d_1, d_2) = \frac{ \langle \phi_i (d_1), S_{i,j} \phi_j (d_2) \rangle }{\|\phi_i(d_1)\| \|\phi_j(d_2)\|},$$
where $d_1$ and $d_2$ are documents written in the $i$-th and $j$-th language respectively. If the the maximal singular value of $S_{i,j}$ is bounded by $1$, then the similarity scores will lie on the interval $[-1, 1]$. In section \ref{sec:models} we will describe some approaches to computing $S_{i,j}$ given training data.

\subsection{Cross-Lingual Models}\label{sec:models}
In this section, we describe several approaches to the problem of computing multilingual simliarities introduced in \ref{sec:tfidf}. We present four approaches:
a simple approach based on $k$-means clustering in \ref{sec:kmeans}, a standard approach based on singular value decomposition in \ref{sec:LSI}, a related
approach called Canonical Correlation Analysis (CCA) in \ref{sec:CCA} and finally a new, more generally applicable method, which is an extension of CCA to more than two languages in \ref{sec:hublang}.

CCA can be used to find
correlated patterns for a pair of languages, whereas the extended method optimizes a
sum of squared correlations between several language pairs. The method takes advantage of the hub language assumption in the training dataset (the English Wikipedia corresponds to a hub) as well as a preprocessing step based on LSI that acts as regularization and improves scalability.

We concentrate on approaches that are based on linear maps rather than alternatives, such as machine translation and probabilistic models, as discussed in the related work section.
We will start by introducing some notation.

\subsection{Notation}

The cross-lingual similarity models presented in this paper are based on comparable corpora. That is, a corpus of documents in multiple languages, with alignment between documents that are of the same topic, or even a rough translation of each other. Wikipedia is an example of a comparable corpus, where a specific entry can be described in multiple languages (e.g. "Berlin" is currently described in 222 languages). News articles represent another example, where the same event can be described by newspapers in several languages.

More formally, \emph{multilingual document} $d = (u_1,\ldots u_m)$ is a tuple of $m$ documents on the same topic (comparable), where $u_i$ is the document written in language $i$. Note that individual document $u_i$ can be an empty document (missing resource) and each $d$ must contain at least two nonempty documents. A comparable corpus $D = {d_1, \ldots, d_N}$ is a collection of multilingual documents. By using the vector space model we can represent $D$ as a set of $m$ matrices $X_1,\ldots,X_m$, where $X_i \in \RR^{n_i \times N}$ is the matrix corresponding to the language $i$ and $n_i$ is the vocabulary size of language $i$. Furthermore, let $X_i^{\ell}$ denote the $\ell$-th column of matrix $X_i$ and the matrices respect the document alignment - the vector $X_i^\ell$ corresponds to the TFIDF vector of the $i$-th component of multilingual document $d_\ell$. We use $N$ to denote the total row dimension of $X$, i.e. $N:= \sum_{i=1}^m n_i$.

\subsection{$k$-means}\label{sec:kmeans}

The $k$-means algorithm is perhaps the most well-known and widely-used clustering algorithm. We now present its application
to computing cross-lingual similarities. The idea is based on concatenating the corpus matrices, running standard $k$-means clustering to obtain the matrix of centroids, "reversing" the concatenation step to obtain a set of aligned bases, which are finally used to compute cross-lingual similarities.

 In order to apply the algorithm, we first merge all the term-document matrices into a single matrix $X$ by stacking the individual term-document matrices:
$$X := \begin{bmatrix}X_1^T ,X_2^T, \cdots, X_m^T \end{bmatrix}^T,$$
such that the columns respect the alignment of the documents (here MATLAB notation for concatenating matrices is used). Therefore, each document  is represented by a long vector indexed by the terms in all languages.

We then run the $k$-means algorithm \cite{kmeans} and obtain a centroid matrix $C \in \RR^{N \times k}$, where the $k$ columns represent centroid vectors. The centroid matrix can be split vertically into $m$ blocks: $$C = [C_1^T \cdots C_m^T]^T,$$ according to the number of dimensions of each language, i.e. $C_i \in \RR^{n_i \times k}$.

The matrices $C_i$ are obtained by analyzing the multilingual corpus matrix $X$ (obtained from Wikipedia for example) and we now show how to use them to compute cross-lingual document similarities on new documents (news documents, for example).

Each matrix $C_i$ represents a vector space basis and can be used to map points in $\RR^{n_i}$ into a $k$-dimensional space, where the coordinates of a vector $x \in \RR^{n_i}$ are expressed as: $$(C_i^T C_i)^{-1} C_i^T x_i.$$

The resulting matrix (when appropriately scaled to have unit norm) for similarity computation between language $i$ and language $j$ is defined as (see the reasoning below):
$$C_i(C_i^T C_i)^{-1} (C_j^T C_j)^{-1} C_j.$$

The matrix is a result of mapping documents in a language independent space using pseudo-inverses of the centroid matrices $P_i = (C_i^T C_i)^{-1} C_i$ and then comparing them using the standard inner product, which results in the matrix $P_i^T P_j$. For the sake of presentation, we assumed that the centroid vectors are linearly independent (an independent subspace could be obtained using an additional Gram-Schmidt step~\cite{golub} on the matrix $C$, if this was not the case).

\subsection{Cross-Lingual Latent Semantic Indexing}\label{sec:LSI}

The next method is Cross-Lingual Latent Semantic Indexing (CL-LSI)\cite{cl_lsi} which is a variant of LSI~\shortcite{lsi} for more than one language. The method is based on computing a truncated singular value decomposition of $X \approx U S V^T$. Since the matrix can be large we can use an iterative method like the Lanczos~\cite{golub} algorithm with reorthogonalization to find the left singular vectors (columns of $U$) corresponding to the largest singular values. It turns out that the Lanczos method converges slowly as the gap between the leading singular values is small. Moreover, the Lanczos method is hard to parallelize. Instead we use a randomized version of the singular value decomposition (SVD) described in \shortcite{tropp} that can be viewed as a block Lanczos method. That enables us to use parallelization and speeds up the computation considerably.

The cross-lingual similarity functions are based on a rank-$k$ truncated SVD: $X \approx U \Sigma V^T,$ where $U \in \RR^{N \times k}$ are basis vectors of interest and $\Sigma \in \RR^{k \times k}$ is truncated diagonal matrix of singular eigenvalues.

An aligned basis is obtained by first splitting $U$ vertically according to the number of dimensions of each language: $U = [U_1^T \cdots U_m^T]^T$. Then, the same as with $k$-means clustering, we compute the pseudoinverses $P_i = (U_i^T U_i)^{-1} U_i^T$. The matrices $P_i$ are used to change the basis from the standard basis in $\RR^{n_i}$ to the basis of columns of $U_i$.

The numerical implementation of the least squares is done by QR algorithm, by computing factorizing $U_i = Q R$, where $Q^TQ = I$ and $R$ is triangular matrix. $P_i$ is then obtained by solving $R P_i = Q$.

\subsection{Canonical Correlation Analysis}\label{sec:CCA}
 The final approach is a statistical technique to analyze data from two sources - after which we will describe our main contribution to computing cross-lingual similarities.

 Canonical Correlation Analysis (CCA)~\cite{Hotelling} is a dimensionality reduction technique similar to Principal Component Analysis (PCA)~\cite{Pearson1901On}, with an additional assumption that the data consists of feature vectors that arose from two sources (two views) that share some information. Examples include: bilingual document collection ~\shortcite{mrpqr} and collection of images and captions~\shortcite{Hardoon_usingimage}. Instead of looking for linear combinations of features that maximize the variance (PCA) we look for a linear combination of feature vectors from the first view and a linear combination for the second view, that are maximally correlated.

Interpreting the columns of $X_i$ as observation vectors sampled from an underlying distribution $\mathcal{X}_i \in \RR^{n_i}$, the idea is to find two weight vectors $w_i \in \RR^{n_i}$ and $w_j \in \RR^{n_j}$ so that the random variables $w_i^T \cdot \mathcal{X}_i$ and $w_j^T \cdot \mathcal{X}_j$ are maximally correlated ($w_i$ and $w_j$ are used to map the random vectors to random variables, by computing weighted sums of vector components). Let $\rho(x,y)$ denote the sample based correlation coefficient between two vectors of observations $x$ and $y$. By using the sample matrix notation $X_i$ and $X_j$ (assuming no data is missing for clearer presentation) this problem can be formulated as the following optimization problem:
\begin{equation*}
\begin{aligned}
& \underset{w_i \in \RR^{n_i}, w_j \in \RR^{n_j}}{\text{maximize}}
& & \rho(w_i^T X_i , w_j^T X_j) = \frac{w_i^T C_{i,j} w_j}{\sqrt{w_i^T C_{i,i} w_i} \sqrt{w_j^T C_{j,j} w_j}},
\end{aligned}
\end{equation*}
where $C_{i,i}$ and $C_{j,j}$ are empirical estimates of variances of $\mathcal{X}_i$ and $\mathcal{X}_j$ respectively and $C_{i,j}$ is an estimate for the covariance matrix. Assuming that the observation vectors are centered (only for the purposes of presentation), the matrices are computed in the following way: $C_{i,j} = \frac{1}{n-1}X_i X_j^T$, and similarly for $C_{i,i}$ and $C_{j,j}$.
The optimization problem can be reduced to an eigenvalue problem and includes inverting the variance matrices $C_{i,i}$ and $C_{j,j}$. If the matrices are not invertible, one can use a regularization technique by replacing $C_{i,i}$ with $(1- \kappa)C_{i,i} + \kappa I$, where $\kappa \in [0,1]$ is the regularization coefficient and $I$ is the identity matrix (the same can be applied to $C_{j,j}$.
A single canonical variable is usually inadequate in representing the original random vector and typically one looks for $k$ projection pairs $(w_i^1, w_j^1),\ldots,(w_i^k, w_j^k)$, so that $(w_i^{u})^T \mathcal{X}_i$ and $(w_j^{u})^T \mathcal{X}_j$ are highly correlated and $(w_i^{u})^T \mathcal{X}_i$ is uncorrelated with $(w_i^{v})^T \mathcal{X}_i$  for $u \neq v$ and analogously for $w_j^u$ vectors.

Note that the method in its original form is only applicable to two languages where an aligned set of observations is available.

\subsection{Hub language based CCA Extension}\label{sec:hublang}
Building cross-lingual similarity models based on comparable corpora is challenging for two main reasons. The problem of missing alignment data: when a number of languages is large, the dataset of documents that cover all languages is small (or may even be empty). Even if only two languages are considered, the set of aligned documents can be small (an extreme example is given by the Piedmontese and Hindi Wikipedias where no inter-language links are available). The second problem is that the number of documents and the involved
1. missing data
2. LSI step
3. Sum of squared correlations

In this section, we describe an extension of CCA to more than two languages which can be trained on large comparable corpora. The main difficulty with applying the
LSI and $k$-means approaches lies in the fact, that when one considers a large number of languages, for example, the top 100 Wikipedia languages (ranked by number of articles), the set of completely
aligned documents is often small or empty. Even if only two languages are considered, the set of aligned documents can be small (for example, Piedmontese and Hindi
Wikipedias had no interlanguage links). In Wikipedia, we observed that even though the training resources are scarce between certain language pairs, there often exists indirect training data
through what we refer to as a hub language.

A \emph{hub language} is a language with a high proportion of non-empty documents in $D = \left\{d_1,..., d_{\ell}\right\}$. The prototypical  exampling  in the case of Wikipedia is English. We use the following notation to define subsets of the multilingual comparable corpus: let $a(i,j)$ denote the index set of all multilingual documents with non-missing data for the $i$-th and $j$-th language:  $$a(i,j) = \left\{k~ |~ d_k = (u_1,...,u_m), u_i \neq \emptyset, u_j \neq \emptyset \right\},$$ and let $a(i)$ denote the index set of all multilingual documents with non missing data for the $i$-th language.

We now describe a two step approach to building a cross-lingual similarity matrix. The first part is related to LSI and reduces the dimensionality of the data. The second step refines the linear mappings and optimizes linear dependence between data.

The first step in our method is to project $X_1, \ldots, X_m$ to lower dimensional spaces without destroying the cross-lingual structure. Treating the nonzero columns of $X_i$ as observation vectors sampled from an underlying distribution $\mathcal{X}_i \in V_i = \RR^{n_i}$, we can analyze the empirical cross-covariance matrices:
$$C_{i,j} = \frac{1}{|a(i,j)|-1 }\sum_{\ell \in a(i,j)} (X_i^{\ell} - c_i)\cdot (X_j^{\ell} - c_j)^T,$$
 where $c_i = \frac{1}{a_i} \sum_{\ell \in a(i)}X_i^{\ell}$. By finding low rank approximations of $C_{i,j}$ we can identify the subspaces of $V_i$ and $V_j$ that are relevant for extracting linear patterns between $\mathcal{X}_i$ and $\mathcal{X}_j$. Let $X_1$ represent the hub language corpus matrix. The LSI approach to finding the subspaces is to perform the singular value decomposition on the full $N \times N$ covariance matrix composed of blocks $C_{i,j}$. If $|a(i,j)|$ is small for many language pairs (as it is in the case of Wikipedia), then many empirical estimates $C_{i,j}$ are unreliable, which can result in overfitting. For this reason we perform the truncated singular value decomposition on the matrix $C = [C_{1,2}  \cdots  C_{1,m}] \approx U S V^T$, where where $U \in \RR^{n_1 \times k}, S \in \RR^{k \times k}, V \in \RR^{(\sum_{i=2}^m n_i) \times k}$. We split the matrix $V$ vertically in blocks with $n_2, \ldots, n_m$ rows: $V = [V_2^T  \cdots  V_m^T]^T$. Note that columns of $U$ are orthogonal but columns in each $V_i$ are not (columns of V are orthogonal). Let $V_1 := U$. We proceed by reducing the dimensionality of each $X_i$ by setting: $Y_i = V_i^T \cdot X_i$, where $Y_i \in \RR^{k\times N}$. To summarize, the first step reduces the dimensionality of the data and is based on CL-LSI, but optimizes only the hub-language related cross-covariance blocks.

The second step involves solving a generalized version of canonical correlation analysis on the matrices $Y_i$ in order to find the mappings $P_i$. The approach is based on the sum of squares of correlations formulation by Kettenring \cite{Kettenring}, where we consider only correlations between pairs $(Y_1, Y_i), i >1$ due to the hub language problem characteristic.
We will present the original unconstrained optimization problem, then a constrained formulation based on the hub language problem characteristic. Then we will simplify the constraints and reformulate
the problem as an eigenvalue problem by using the method of Lagrange multipliers.

The original sum of squared correlation is formulated as an unconstrained problem:
\begin{equation*}
  \begin{aligned}
    & \underset{w_i \in \RR^{k}}{\text{maximize}}
    & & \sum_{i < j}^m  \rho(w_i^T Y_i, w_j^T Y_j)^2.
\end{aligned}
\end{equation*}
We solve a similar problem by restricting $i=1$ and omit optimizing over non-hub language pairs.
Let $D_{i,i} \in \RR^{k \times k}$ denote the empirical covariance of $\mathcal{Y}_i$ and $D_{i,j}$ denote the empirical cross-covariance computed based on $\mathcal{Y}_i$ and $\mathcal{Y}_j$. We solve the following constrained (unit variance constraints) optimization problem:
\begin{equation}\label{squaredCorHubOriginal}
  \begin{aligned}
    & \underset{w_i \in \RR^{k}}{\text{maximize}}
    & & \sum_{i = 2}^m  \left(w_1^T D_{1,i} w_i \right)^2
    & \text{subject to}
    & & w_i^T D_{i,i} w_i = 1, \quad\forall i = 1,\ldots, m.
\end{aligned}
\end{equation}
The constraints $w_i^T D_{i,i} w_i$ can be simplified by using the Cholesky decomposition $D_{i,i} = K_i^T \cdot K_i$ and substitution: $y_i := K_i w_i$. By inverting the $K_i$ matrices and defining  $G_i := K_1^{-T} D_{1,i} K_i^{-1}$, the problem can be reformulated:
\begin{equation}\label{squaredCorHub}
  \begin{aligned}
    & \underset{y_i \in \RR^{k}}{\text{maximize}}
    & & \sum_{i = 2}^m  \left(y_1^T G_{i} y_i \right)^2
    & \text{subject to}
    & & y_i^T y_i = 1, \quad\forall i = 1,\ldots, m.
\end{aligned}
\end{equation}
A necessary condition for optimality is that the derivatives of the Lagrangian vanish. The Lagrangian of (\ref{squaredCorHub}) is expressed as:
$$  L(y_1, \ldots, y_m, \lambda_1, \ldots, \lambda_m) = \sum_{i = 2}^m  \left(y_1^T G_{i} y_i \right)^2 + \sum_{i=1}^m \lambda_i \left(y_i^T y_i - 1\right).$$
Stationarity conditions give us:
\begin{equation}\label{dLdx1}
 \frac{\partial}{\partial x_1} L = 0 \Rightarrow \sum_{i = 2}^m  \left(y_1^T G_{i} y_i \right) G_i y_i + \lambda_1 y_1,
\end{equation}
\begin{equation}\label{dLdxi}
\frac{\partial}{\partial x_i} L = 0 \Rightarrow \left(y_1^T G_{i} y_i \right) G_{i}^T y_1 + \lambda_i y_i = 0,~i > 1.
\end{equation}
Multiplying the equations (\ref{dLdxi}) with $y_i^T$ and applying the constraints, we can eliminate $\lambda_i$ which gives us:
\begin{equation}\label{eqy1yi}
G_{i}^T y_1 = \left(y_1^T G_{i} y_i \right) y_i,~i > 1.
\end{equation}
Plugging this into (\ref{dLdx1}), we obtain an eigenvalue problem:
$$\left( \sum_{i = 2}^m G_i G_{i}^T \right) y_1 + \lambda_1 y_1 = 0.$$
The eigenvectors of $\left( \sum_{i = 2}^m G_i G_{i}^T \right)$ solve the problem for the first language. The solutions for $y_i$ are obtained from (\ref{eqy1yi}): $y_i := \frac{G_{i}^T y_1}{\| G_{i}^T y_1 \|}$.
Note that the solution (\ref{squaredCorHubOriginal}) to can be recovered by: $w_i := K_i^{-1} y_i$. The linear transformation of the $w$ variables are thus expressed as:
$$ Y_1 := \text{eigenvectors of} \sum_{i = 2}^m G_i G_{i}^T, $$
$$ W_1 = K_1^{-1} Y_1 $$
$$ W_i = K_i^{-1} G_{i}^T Y_1 N,$$
where $N$ is a diagonal matrix that normalizes $G_{i}^T Y_1$, with $N(j,j) := \frac{1}{\|G(_{i} Y_1(:,j)\|}$.

 The technique is related to the Generalization of Canonical Correlation Analysis (GCCA) by Carroll \citeyear{Carroll}, where an unknown group configuration variable is defined and objective is to maximize the sum of squared correlation between the group variable and the others. The problem can be reformulated as an eigenvalue problem. The difference lies in the fact that we set the unknown group configuration variable as the hub language, which simplifies the solution. The complexity of our method is $O(k^3)$, whereas solving the GCCA method scales as $O(N^3)$ where $N$ is the number of samples (see \cite{gifi}). Another issue with GCCA is that it cannot be directly applied to the case of missing documents.

To summarize, we first reduced the dimensionality of our data to $k$-dimensional features and then found a new representation (via linear transformation) that maximizes directions of linear dependence between the languages. The final projections that enable mappings to a common space are defined as: $P_i(x) = W_i^T V_i^T x.$

\section{Cross-lingual Event Linking}\label{sec:linking}

The main application on which we test the above similarity is cross-lingual event linking. In online media streams -- particularly new articles -- there is often duplication of reporting, different viewpoints or opinions, all centering around a single event. Same events are covered by many articles and the question we address is how to find all the articles in different languages that are describing a single event. We base our evaluation on an online system for detection of world events, called Event Registry. We do not address the problem of detection of events in this paper but rather consider the problem that given a series of events, how may we best ``tag" or match articles in different languages to these events. The events are represented by clusters of articles and so ultimately our problem reduces to finding suitable matchings between clusters with articles in different languages.

\subsection{Problem definition}

The problem of cross-lingual event linking is to identify the clusters describing the same events. Each article $a_i$ is written in a language $l$, where $\ell \in L = \{\ell_1,\ell_2,...,\ell_k\}$. For each language $\ell$, we generate and maintain a set of language-specific clusters $C_l$. That is, all articles are in language $l$. For any cluster $c_i \in C_{\ell_a}$, we would like to identify clusters $c_j \in C_{\ell_b}$ (where $\ell_a \neq \ell_b$) that describe the same event as $c_i$. We define \emph{equivalent clusters} as  pairs $(c_i,c_j)$ such that both clusters in the pair describe the same event in languages $C_{\ell_a}$ and $C_{\ell_b}$ respectively. Furthermore, we iterate over all the languages we consider. That is, given a cluster in language $\ell_i$, we look for matching clusters in language $\ell_j$ for $i\neq j$  in a pairwise fashion.

\begin{figure}[tb]
\centering
\includegraphics[width=0.6\textwidth]{clusters}
\caption{\label{fig:clusters}  Given clusters of articles in different languages, the goal of  linking is to match clusters which have similar content across languages. The challenge in linking is to overcome the errors in the clustering and differing content as well as make the approach scalable. Therefore, we first identify candidate clusters. }
\end{figure}

Furthermore, the matching of clusters is a \emph{generalized matching}. We cannot assume that there is only one cluster per language per event, nor can we assume complete coverage -- that there exists at least one cluster per event in every language. This implies that we cannot make any assumptions on the matching, e.g. one-to-one, complete, etc. This excludes the use of standard weighted bipartite matching type of algortihms for this problem. An example is shown in Figure~\ref{fig:clusters}, where a cluster may contain articles which are closely matched with many clusters in a different language.

We impose an additional requirement.  Identification of the equivalent clusters cannot be done exhaustively -- for a cluster $c_i$ we wish to avoid testing all other clusters for all the other languages. Performing exhaustive comparison for each cluster $c_i$ would result in $O(N^2)$ tests, where $N$ is the number of all clusters (over all languages), which is not feasible when the number of clusters is on the order of tens of thousands.

\begin{comment}
One of the use cases where article linking information can be applied is also in an online system for detection of world events, called Event Registry\cite{}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles. The term event is vague and ambiguous, but for the practical purposes we define it as ``any significant happening that is being reported about in the media''.

Event Registry works by collecting and analyzing news articles that are published by news outlets in different languages all over the world. Collected articles are first semantically annotated by identifying in them mentions of relevant concepts -- either entities or important keywords. The disambiguated and entity linking of the concepts is done using Wikipedia as the main knowledge base. The key step in Event Registry is then to apply an online clustering algorithm on the articles in order to identify groups of articles that are discussing the same event. For each new article, the clustering algorithm determines if the article should be assigned to some existing cluster or into a new cluster. The underlying assumption is that articles that are describing the same event are similar enough and will therefore be put into the same cluster.

In short, the article clustering works as follows (details are available in \cite{Brank2014}). First we create a new document by joining the title and the body of the article. We then tokenize the document, remove the stop words and stem the words. The remaining words are represented as a vector in the vector-space model. Each value in the vector is additionally normalized using TF-IDF. After normalization, we use the vector to find the most similar existing cluster. To be able to compare the vector with a cluster we first need a vector representation for each of the clusters. This is done by computing a centroid vector of the cluster by simply averaging the vectors of the articles in the cluster. By computing the cosine similarity of the article vector with each of the centroid vectors we then find the cluster with the highest similarity. If the highest similarity is above the selected threshold (we used 0.4 in our experiments), the article is assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article.

Since articles about an event are commonly written only for a short period of time, we remove clusters once the oldest article in the cluster becomes more than 4 days old. This housekeeping mechanism prevents the clustering from becoming slow and also ensures that articles are not assigned to obsolete clusters.

In order to treat the cluster as describing an event in Event Registry, the cluster has to grow above a certain number of articles (the number depends on the language). This condition ensures that only content that receives a sufficient coverage in the media is treated as an event. Once a new event is identified, a new unique ID is assigned to it and the main information about the event is then automatically extracted by analyzing the articles assigned to it. The extracted information includes properties such as the date of the event, the location, who is involved in it, what the event is about, etc.

The clustering of articles is done for each language separately -- English articles are, for example, clustered separately from Spanish articles. When events are identified, they will therefore be identified from clusters in each language separately. As a result, the same event will be represented in Event Registry with two or more event IDs, and each event will only contain coverage in a single language. To avoid this problem we want to exploit the available information about the similarity of the articles in a cluster with articles in other languages. Using this information we hope to be able to connect clusters in different languages and represent them using a single event ID.

\subsection{Problem definition}

The problem in the second phase of the pipeline is to identify the clusters (given?) describing the same events. Each article $a_i$ is written in a language $l$, where $l \in L = \{l_1,l_2,...,l_k\}$. For each language $l$, we generate and maintain a set of language-specific clusters $C_l$. That is, all articles are in language $l$. For any cluster $c_i \in C_{l_a}$, we would like to identify clusters $c_j \in C_{l_b}$ (where $l_a \neq l_b$) that describe the same event as $c_i$. We define \emph{equivalent clusters} as  pairs $(c_i,c_j)$ such that both clusters in the pair describe the same event in languages $C_{l_a}$ and $C_{l_b}$ respectively.

\end{comment}

\subsection{Algorithm}

In order to identify clusters $c_j$ that are equivalent to cluster $c_i$, we have developed a two-stage algoirhtm. For a cluster $c_i$, we first efficiently identify a small set of candidate clusters and then find those clusters among the candidates, which are equivalent to $c_i$. An example is  shown in  Figure~\ref{fig:clusters}.

The details of the first step are described in Algorithm~\ref{cluster_merge_algo1}. The algorithm begins by individually inspecting each article $a_i$ in the cluster $c_i$. Using a chosen method for computing cross-lingual document similarity (see Section~\ref{sec:models}), it identifies the 10 most similar articles to $a_i$ in each language $l \in L$. For each  similar article $a_j$, we identify its corresponding  cluster $c_j$ and add it to the set of candidates. The set of candidate clusters obtained in this way is several orders of magnitude smaller than the number of all clusters. Although computed document similarities are approximate, our  assumption  is that articles in different languages describing the same event, will generally have a higher similarity than articles about different events. While this assumption does not always hold, redundancy in the data mitigates these false positives. Since we compute the 10 most similar articles for each article in $c_i$, we are likely to identify all the relevant candidates for cluster $c_i$.

\begin{algorithm}[t!]
\SetKwInput{KwInput}{input}\SetKwInput{KwOutput}{output}
\KwInput{test cluster $c_i$,\\a set of clusters $C_l$ for each language $l \in L$}
\KwOutput{a set of clusters $C$ that are potentially equivalent to $c_i$}
$C = \{\}$\;
\For{article $a_i \in c_i$} {
    \For{language $l \in L$} {
        \tcc{use hub CCA to find 10 most similar articles to article $a_i$ in language $l$}
        $SimArt = getCCASimArts(a_i, l)$\;
        \For{article $a_j \in SimArt$} {
            \tcc{find cluster $c_j$ to which article $a_j$ is assigned to}
            $c_j := c ; c \in C_l, a_j \in c$\;
            \tcc{add cluster $c_j$ to the set of candidates $C$}
            $C \leftarrow c_j$\;
        }
    }
}
\caption{Algorithm for identifying candidate clusters $C$ that are potentially equivalent to $c_i$}
\label{cluster_merge_algo1}
\end{algorithm}

The second stage of the algorithm (described in the Algorithm~\ref{cluster_merge_algo2}.) determines which (if any) of the candidate clusters are equivalent to $c_i$. We treat this task as a supervised learning problem. For each candidate cluster $c_j \in C'$, we compute a vector of learning features that should be indicative of whether the $c_i$ and $c_j$ are equivalent or not. The details of our chosen features will be described in the following section. Once these features are computed, we apply a classification model which was trained using a  dataset  of hand-labeled examples. If the model predicts that tested clusters are equivalent, the $c_j$ is added to the set $C'$.

\begin{algorithm}[tb!]
\SetKwInput{KwInput}{input}\SetKwInput{KwOutput}{output}
\KwInput{test cluster $c_i$,\\
a set of candidate clusters $C$ that are potentially equivalent to $c_i$\\
classification model $M$ trained on the hand-labeled examples}
\KwOutput{a set of clusters $C'$ that are equivalent to $c_i$}

$C' = \{\}$\;
\For{cluster $c_j \in C$} {
    \tcc{extract a vector of features for cluster pair $c_i$ and $c_j$}
    $\vec{v} = f(c_i, c_j)$\;
    \tcc{classify the feature vector $\vec{v}$ using the model $M$}
    $pred = M(\vec{v})$\;
    \If{$pred$ = true} {
        $C' \leftarrow c_j$
    }
}
\caption{Algorithm for identifying clusters $C'$ that are equivalent to cluster $c_i$}
\label{cluster_merge_algo2}
\end{algorithm}

\subsection{Training a classification model}

In order to train a classification model for predicting whether two clusters are equivalent, we must first have labelled data. To obtain the data, we used two human annotators. The annotators were provided with an interface displaying information about the articles in two clusters and their task was to determine if the clusters were equivalent or not. To obtain a pair of clusters $(c_i, c_j)$ to annotate, we first randomly chose a cluster $c_i$, used the Algorithm~\ref{cluster_merge_algo1} to compute a set of potentially equivalent clusters $C$ and randomly chose a cluster $c_j \in C$.

The dataset provided by the annotators contains 808 examples, of which 402 are equivalent clusters pairs and 406 are not. Clusters in each learning example are either in English, Spanish or German. Although Event Registry imports articles in other languages as well, we restricted our experiments to these three languages. We chose only these three languages since they have very large number of articles and clusters per day which makes the cluster linking problem hard due to large number of possible links.

To build a classification model we first extract  informative features from the labeled examples. Based on the availability of the data, we extracted the following groups of features:

\begin{itemize}
\item {\bf Cross-lingual article linking features}. Cross-lingual article linking information was already used to obtain candidates for the equivalent clusters. Additionally, it can also be used to provide valuable  features for learning. There are two main features that we extract from article links between clusters -- \texttt{linkCount} and \texttt{avgSimScore}. \texttt{linkCount} is the number of times an article $a_i \in c_i$ has as one of the 10 most similar articles an article $a_j \in c_j$. In other words, it is the number of times an article from $c_i$ has a very similar article, which is in $c_j$. Beside the number of links between the two clusters we can also take into account the similarity scores of the links. The \texttt{avgSimScore} is the feature that represents the average similarity score of the links between the two clusters.

\item {\bf Concept related features}. Articles that are imported into Event Registry are first semantically annotated by linking mentioned entities and keywords to the corresponding Wikipedia pages. Whenever Barack Obama is, for example, mentioned in the article, the article is annotated with a link to his Wikipedia page\footnote{\emph{http://en.wikipedia.org/wiki/Barack\_Obama}}. In the same way all mentions of people, locations, organizations and even ordinary keywords (e.g. bank, tax, ebola, plane, company) are annotated. Although the Spanish article about Obama will be annotated with his Spanish version of the Wikipedia page, in many cases we can link the Wikipedia pages to their English versions. This can be done since Wikipedia itself provides information regarding which pages in different languages represent the same concept/entity. Using this approach, the word ``avi\'on'' in a Spanish article will be annotated with the same concept as the word ``plane'' in an English article. Although the articles are in different languages, the annotations can therefore provide a language-independent vocabulary that can be used to compare articles/clusters. By analyzing all the articles in clusters $c_i$ and $c_j$, we can identify the most relevant entities and keywords for each cluster. Additionally, we can also assign weights to the concepts based on how frequently they occur in the articles in the cluster. From the list of relevant concepts and corresponding weights, we consider the following features: \texttt{entityCosSim} which represents the cosine similarity on the vector of entities mentioned in both clusters; \texttt{keywordCosSim} which represents the cosine similarity on the keywords only; for cases where the weights are not relevant, we have also added two additional features - (\texttt{entityJaccardSim} and \texttt{keywordJaccardSim}), where Jaccard similarity is used instead of the cosine similarity.

\item {\bf Miscellaneous features}. This group contains three miscellaneous features that seem discriminative but are unrelated to the previous two groups. The first feature is related to the event location. By considering the locations mentioned in the articles in a cluster, we are able to estimate where an event occurred. The \texttt{hasSameLocation} feature is a boolean variable that is true when the location of the event in both clusters is the same. Another important feature is related to the time when the articles in the clusters were published. For each cluster, we first compute the average publication time and date of the articles. The \texttt{timeDiff} is then defined as the absolute difference in hours between the two computed dates. The final feature we consider is computed by analyzing the mentions of dates in the articles. Using an extensive set of regular expressions, we are able to detect mentions of dates in different forms. To compute the feature \texttt{sharedDates} we start by identifying the list of dates mentioned in articles in each cluster separately. The value of \texttt{sharedDates} is then determined as the Jaccard similarity on the two lists of dates.

\end{itemize}

Using the  features described above and the user-provided labels, we train a classification model $M$ which is used in Algorithm~\ref{cluster_merge_algo2} to classify a new cluster pair as equivalent or not. The classification algorithm that we used to train a model was a linear Support Vector Machine (SVM) method \shortcite{shawe-taylor04kernel}.

\section{Evaluation}

We will describe the main dataset for building cross-lingual models which is based on Wikipedia. After that we will present two sets of experiments. The first set of experiments
will establish that the hub based approach can deal with language pairs where little or no training data is available. The second set of experiments will compare the main approaches
that we presented on the task of mate retrieval and the task of event linking. We will also examine how different choices of features impact the event linking performance.

\subsection{Wikipedia Comparable Corpus}

To investigate the empirical performance of the low rank approximations we will test the algorithms on a large-scale, real-world multi-lingual dataset that we extracted from Wikipedia by using inter-language links for alignment. This  results in a large number of weakly comparable documents in more than $200$ languages. Wikipedia is a large source of multi-lingual data that is especially important for the languages for which no translation tools, multi-lingual dictionaries as Eurovoc \cite{eurovoc}, or strongly aligned multi-lingual corpora as Europarl \cite{europarl} are available. Documents in different languages are related with so called 'inter-language' links that can be found on the left of the Wikipedia page. The Wikipedia is constantly growing. There are currently 12 Wikipedias with more than 1 million %$10^6$
 articles, $52$ with more than 100k %$10^5$
 articles, $129$ with more than 10k articles, and $236$ with more than $1000$ articles.

Each Wikipedia page is embedded in the page tag. First, we check if the title of the page consists of any special namespace and do not process such pages. Then, we check if this is a redirection page and we store the redirect link as inter-language links can point to redirection link also. If none of the above applies, we extract the text and parse the Wikipedia markup. Currently, all the markup is removed.

We get inter-language link matrix using previously stored redirection links and inter-language links. If inter-language link points to the redirection we replace it with the redirection target link. It turns out that we obtain the matrix $M$ that is not symmetric, consequently the underlying graph is not symmetric. That means that existence of the inter-language link in one way (i.e. English to German) does not guarantee that there is an inter-language link in the reverse direction (German to English). To correct this we transform this matrix to symmetric by computing $M+M^T$ and obtaining an undirected graph. In the rare case that after symmetrization we have multiple links pointing from the document, we pick the first one that we encountered. This matrix enables us to build an alignment across all Wikipedia \footnote{https://www.wikipedia.org/ dumps available in 2013} languages.

\subsection{Experiments With Missing Alignment Data}\label{experiments:hubcca}

 In this subsection, we will investigate the empirical performance of hub CCA approach. We will demonstrate that this approach can be successfully applied even in the case of fully missing alignment information.
 To this purpose, we select a subset of Wikipedia languages containing three major languages, English (4,212k articles)--\emph{en} (hub language), Spanish (9,686k articles)--\emph{es}, Russian (9,662k articles)--\emph{ru}, and five minority (in the sense of Wikipedia sizes) languages, Slovenian (136k articles)--\emph{sl}, Piedmontese (59k articles)--\emph{pms}, Waray-Waray (112k articles)--\emph{war} (all with about 2 million native speakers), Creole (54k articles)--\emph{ht} (8 million native speakers), and Hindi (97k articles)--\emph{hi} (180 million native speakers). For preprocessing, we remove the documents that contain less than 20 different words (stubs \footnote{This documents are typically garbage, the titles of the columns in the table, remains of the parsing process, or Wikipedia articles with very little or no information contained in one or two sentences.}) and remove words occurring in less than 50 documents as well as the top 100 most frequent words (in each language separately). We represent the documents as normalized TFIDF\cite{Salton88term-weightingapproaches} weighted vectors.
 %These are languages with Wikipedia article sizes comparable to the minority languages.  (since the number of speakers does not directly correlate with the number of Wikipedia articles). The prime hub candidate is the English %language which is well aligned with all other Wikipedia languages, although the alignment quality varies quite a bit. Furthermore, we remove the documents that contain less than 20 different words and remove words that are too %infrequent as well as the top 100 most frequent words in the vocabularies.
 Although the English language is well aligned, number of articles highly correlates with the
    number of articles connected to English, with all Wikipedia languages, we must note that quality of alignment varies quite a bit, especially in the case of small Wikipedias.
% \todo{Move this to footnote for stubs}

The evaluation is based on splitting the data into training and test sets. %(which are described later).
We select the test set documents as all multi-lingual documents with at least one nonempty alignment from the list: (\emph{hi}, \emph{ht}), (\emph{hi}, \emph{pms}), (\emph{war}, \emph{ht}), (\emph{war}, \emph{pms}). This guarantees that we cover all the languages. Moreover this test set is suitable for testing the retrieval thorough the hub as the chosen pairs have empty alignments. The remaining documents are used for training. In Table \ref{table:train_test}, we display the corresponding sizes of training and test documents for each language pair.

On the training set, we perform the two step procedure to obtain the common document representation as a set of mappings $P_i$. A test set for each language pair, $test_{i,j} = \{(x_\ell,y_\ell) | \ell = 1:n(i,j)\} $, consists of comparable document pairs (linked Wikipedia pages), where $n(i,j)$ is the test set size. We evaluate the representation by measuring mate retrieval quality on the test sets: for each $\ell$, we rank the projected documents $P_j(y_1),\ldots, P_j(y_{n(i,j)})$ according to their similarity with $P_i(x_\ell)$ and compute the rank of the mate document $r(\ell) = rank(P_j(y_\ell))$. The final retrieval score (between -100 and 100) is computed as: $\frac{100}{n(i,j)} \cdot \sum_{\ell = 1}^{n(i,j)} \left( \frac{n(i,j) - r(\ell)}{n(i,j) -1} -0.5\right)$. A score that is less than 0 means that the method performs worse than random retrieval and a score of 100 indicates perfect mate retrieval. The mate retrieval results are included in Table \ref{table:retrieval}.

We observe that the method performs well on all pairs between languages: \emph{en}, \emph{es}, \emph{ru}, \emph{sl}, where at least 50,000 training documents are available. We notice that taking $k = 500$ or $k = 1000$ multilingual topics usually results in similar performance, with some notable exceptions: in the case of (\emph{ht}, \emph{war}) the additional topics result in an increase in performance, as opposed to (\emph{ht},\emph{pms}) where performance drops, which suggests overfitting. The languages where the method performs poorly are \emph{ht} and \emph{war}, which can be explained by the quality of data (see Table \ref{table:rank} and explanation that follows). In case of \emph{pms}, we demonstrate that solid performance can be achieved for language pairs (\emph{pms}, \emph{sl}) and (\emph{pms}, \emph{hi}), where only 2000 training documents are shared between \emph{pms} and \emph{sl} and no training documents are available between \emph{pms} and \emph{hi}. Also observe that in the case of (\emph{pms}, \emph{ht}) the method still obtains a score of 62, even though training set intersection is zero and \emph{ht} data is corrupted, which we will show in the next paragraph.
{
\renewcommand\tabcolsep{3pt}
\begin{table}[h!]
\caption{Pairwise retrieval, 500 topics on the left -- 1000 topics on the right}\label{table:retrieval}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
\cline{1-9}
&	en&	es&	ru&	sl&	hi&	war&	ht&	pms\\\cline{1-9}
en&	    &	98~-~98&	95~-~97&	97~-~98&	82~-~84&	76~-~74&	53~-~55&	 96~-~97\\
\cline{1-9}
es&	97~-~98&	&	94~-~96&	97~-~98&	85~-~84&	76~-~77&	56~-~57&	96~-~96\\
\cline{1-9}
ru&	96~-~97&	94~-~95&	&	97~-~97&	81~-~82&	73~-~74&	55~-~56&	96~-~96\\
\cline{1-9}
sl&	96~-~97&	95~-~95&	95~-~95&	&	91~-~91&	68~-~68&	59~-~69&	93~-~93\\
\cline{1-9}
hi&	81~-~82&	82~-~81&	80~-~80&	91~-~91&	&	68~-~67&	50~-~55&	87~-~86\\
\cline{1-9}
war&	68~-~63&	71~-~68&	72~-~71&	68~-~68&	66~-~62&	&	28~-~48&	 24~-~21\\
\cline{1-9}
ht&	52~-~58&	63~-~66&	66~-~62&	61~-~71&	44~-~55&	16~-~50&	&	62~-~49\\
\cline{1-9}
pms&	95~-~96&	96~-~96&	94~-~94&	93~-~93&	85~-~85&	23~-~26&	66~-~54&	 \\
\cline{1-9}
\end{tabular}
\end{center}
\end{table}
}
%
%We now describe the selection of train and test sets. We select the test set documents as all multi-lingual documents with at least one nonempty alignment from the list: (\emph{hi}, \emph{ht}), (\emph{hi}, \emph{pms}), (\emph{war}, \emph{ht}), (\emph{war}, \emph{pms}). This guarantees that we cover all the languages. Moreover this test set is suitable for testing the retrieval thorough the hub as the chosen pairs have empty alignments. The remaining documents are used for training. In Table \ref{table:train_test}, we display the corresponding sizes of training and test documents for each language pair. The first row represents the size of the training sets used to construct the mappings in low dimensional language independent space using the English--\emph{en} as a hub. The diagonal elements represent number of the unique training documents and test documents in each language.

{
\renewcommand\tabcolsep{3pt}
\begin{table}[h!]
\centering
\caption{Pairwise training -- test sizes (in thousands)
The first row represents the size of the training sets used to construct the mappings in low dimensional language independent space using the English--\emph{en} as a hub. The diagonal elements represent number of the unique training documents and test documents in each language.
}
\label{table:train_test}
{
\small
\begin{tabular}{c|c|c|c|c|c|c|c|c|}
&	en&	es&	ru&	sl&	hi&	war&	ht&	pms\\\cline{1-9}
en&	671~-~4.64&	463~-~4.29&	369~-~3.19&	50.3~-~2&	14.4~-~2.76&	8.58~-~2.41&	 17~-~2.32&	 16.6~-~2.67\\
\cline{2-9}
es&	\multicolumn{1}{c|}{}	&	463~-~4.29&	187~-~2.94&	28.2~-~1.96&	8.72~-~2.48&	 6.88~-~2.4&	 13.2~-~2&	 13.8~-~2.58\\
\cline{3-9}
ru&	\multicolumn{2}{c|}{}	&	369~-~3.19&	29.6~-~1.92&	9.16~-~2.68&	2.92~-~1.1&	 3.23~-~2.2&	 10.2~-~1.29\\
\cline{4-9}
sl&	\multicolumn{3}{c|}{}	&	50.3~-~2&	3.83~-~1.65&	1.23~-~0.986&	0.949~-~1.23&	 1.85~-~0.988\\
\cline{5-9}
hi&	\multicolumn{4}{c|}{}	&	14.4~-~2.76&	0.579~-~0.76&	0.0~-~2.08&	0.0~-~0.796\\
\cline{6-9}
war&	\multicolumn{5}{c|}{}	&	8.58~-~2.41&	0.043~-~0.534&	0.0~-~1.97\\
\cline{7-9}
ht&	\multicolumn{6}{c|}{}	&	17~-~2.32&	0.0~-~0.355\\
\cline{8-9}
pms&	\multicolumn{7}{c|}{}	&	16.6~-~2.67\\
\cline{9-9}
\end{tabular}
}
\end{table}
}

We further inspect the properties of the training sets by roughly estimating the fraction \texttt{rank(A)/min(size(A))} for each training English matrix and its corresponding mate matrix. Ideally, these two fractions are approximately the same so  both aligned spaces should have reasonably similar dimensionality. We display these numbers as pairs in Table \ref{table:rank}.

\begin{table}[h]
\caption{Dimensionality drift}
\label{table:rank}
\begin{tabular}{|c|c|c|c|c|c|c|}
\cline{1-7}
(en, de)     &   (en, ru)     &   (en, sl)       &     (en, hi)&   (en, war)      &      (en, ht) &   (en, pms)\\
\cline{1-7}
(0.81, 0.89)   &  (0.8, 0.89)  &   (0.98, 0.96)    &    (1, 1)  &   (0.74, 0.56)  &      (1, 0.22)  &   (0.89, 0.38)\\
\cline{1-7}
\end{tabular}
\end{table}

It is clear that in the case of Creole language only at most $22\%$ documents are unique and suitable for the training. Though we removed the stub documents, many of remaining documents are nearly the same, as the quality of some minor Wikipedias is low. This was confirmed for Creole, Waray-Waray, and Piedmontese language by manual inspection. The low quality documents correspond to templates about the year, person, town, etc. and contain very few unique words.

There is also a problem with the quality of the test data. For example, if we look at test pair (\emph{war}, \emph{ht}) only 386/534 Waray-Waray test documents are unique but on other side almost all Creole test documents (523/534) are unique. This indicates a poor alignment which leads to poor performance.
%}

\subsection{Evaluation Of Cross-Lingual Event Linking}
In order to determine how accurately we can predict cluster equivalence, we performed two experiments in multi-lingual setting, English, German and Spanish language for which we had labelled data to evaluate the linking performance. In the first experiment, we tested how well  the individual approaches for cross-lingual article linking perform when used for linking the clusters about the same event. In the second experiment we tested how accurate is the prediction model when trained on different subsets of learning features. To evaluate the prediction accuracy for a given dataset we used 10-fold cross validation.
%In this technique, the dataset is partitioned into 10 folds, where 9 folds are used for training a model and the left-out fold is then used to test the accuracy of the model. The learning %and testing process is repeated 10 times, each time with a different left-out fold.

% which method?
In Section~\ref{sec:models}, we  described three main algorithms for identifying similar articles in different languages. These algorithms were $k$-means, LSI and hub CCA. As a training set, we used common Wikipedia alignment for all three languages. To test which of these algorithms performed best, we performed the following test. Using each algorithm, we analyzed all articles in Event Registry to find the most similar articles in other languages. The values of the cross-lingual article linking features (\texttt{linkCount} and \texttt{avgSimScore}) were then recomputed for all annotated learning examples. A dataset with only these two features and the label was constructed and used to test how accurate prediction models can be induced using the data. The results of the experiment are shown in Table~\ref{table:linkingEvalAlgos}.
%The table shows for each of the algorithms the obtained classification accuracy, precision and recall.
 We can see that the most informative features for cluster linking are obtained using the hub CCA algorithm, while $k$-means performs the worse.

% AMMR experiments
We also compared the proposed approaches on the task of Wikipedia mate retrieval (the same task as in Section~\ref{experiments:hubcca}). We computed the Average (over language pairs) Mean Reciprocal Rank (AMRR) \shortcite{voorhees1999trec}  performance of the different approaches on the  Wikipedia data by holding out $15,000$ aligned test documents and using $300,000$ aligned documents as the training set. The Figure~\ref{pic:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that hub CCA outperforms LSI approach and K-means lags far behind when testing on Wikipedia data. The hub CCA approach with $500$ features manages to perform comparably to $LSI$ based approach with $1000$-features, which shows that the $CCA$ method can improve both model memory footprint as well as similarity computation time.

% number of topics?
Furthermore, we inspected how the number of features (topics) influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing number of feature vectors larger than $500$ barely affects linking performance, which is in contrast with the fact that additional features helped improve AMMR, see Figure~\ref{pic:AMRR}. Such differences may have arisen due to different domains of training and testing (Wikipedia pages versus news articles).

% cluster size?
We also analyzed how cluster size influences the linking performance. The intuition is that clusters with large number of documents are easier to link, which can be seen in Table~\ref{table:linkingEvalAlgosLargeSmall}. Large clusters aggregate a lot of information, so the performance of the similarity model is not as important so long as we assign articles to correct clusters. In the case of small clusters, we expect the accuracy to matter more. Our experiments indicate that this is indeed the case: it is harder to link small clusters and better retrieval accuracy may matter more. Hub CCA outperforms other methods in this sense in terms of precision, which is consistent with its AMRR performance, see Figure~\ref{pic:AMRR}.
It is worth noting that added accuracy is also beneficial when we wish to choose good representatives of clusters. Therefore, we decided to use hub CCA in the cluster linking component.

%{\oldText{TODO:} add some explanation why this might be so.}
%\emph{Andrej adds some tests, new experiment description graphs, and explanations for performance, done by Monday}
%15000 test documents, around 300 000 sth training documents
%Results for feature spaces of dimensions 500, 800, 1000  overfitting, 800 minor improvement
%Timex complexity of similarity computation scales linearly with dimension space. We double numbers of articles and decrease memory requirements by 1/2.
%CCA       CA          AUC         Prec        Recall
%CCA =     [0.7882      0.8504      0.7692	   0.8209
%           0.7957      0.8624      0.7788      0.8234
%           0.8031      0.8654      0.8015      0.8035];
%LSIS=     [0.7994      0.8627      0.7778      0.8358
%           0.7956      0.8634      0.7828      0.8159
%           0.7994      0.8642      0.8030      0.7910];
%LSIW=     [0.7894      0.8606      0.7673      0.8284;
%           0.7870      0.8632      0.7662      0.8234;
%           0.8018      0.8663      0.7854      0.8284];

%AMPR as function of the number of feature vectors
%500 is good, 800 better, 1000 overfitting, noise!
\begin{figure}
%\includegraphics{slika.png}
\centering
\input{retrieval}
\caption{Average of mean reciprocal ranks}
\label{pic:AMRR}
\end{figure}
%Why the hub approach?
%Optimize for language pairs in the system that have the largest number of articles in the system!
%Is star CCA really the best???
%Maybe omit this table?
\begin{table}[h]
\caption{Accuracy of cluster linking with 500 - 800 - 1000 topic vectors obtained from different cross-lingual similarity algorithms. The table shows for each of the algorithms the obtained classification accuracy, precision and recall.}
\label{table:linkingEvalAlgos}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Models & A & Precision & Recall \\ \cline{1-4}
  CCA      & 78.2 - 79.6 - 80.3 \% & 76.3 - 78.0 - 80.5 \%  & 81.6 - 82.1 - 79.9 \%  \\ \cline{1-4}
  LSI      & 78.9 - 78.7 - 80.6 \%  & 76.8 - 77.0 - 78.7 \%  & 83.3 - 80.6 - 83.6 \%  \\ \cline{1-4}
 $k$-means & 73.9\phantom{ - 78.7 - 80.6 }\% & 69.5\phantom{ - 78.7 - 80.6 }\%  & 84.6\phantom{ - 78.7 - 80.6 }\%  \\ \cline{1-4}
\end{tabular}
\end{center}
\end{table}

\begin{table}[h]
\caption{Accuracy of cluster linking using $500$ features on two datasets containing large (left number) and small (right number) clusters}
\label{table:linkingEvalAlgosLargeSmall}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Models & CA & Precision & Recall \\ \cline{1-4}
  CCA      & 81.2 - 77.8 \% & 80.5 - 74.5 \%  & 91.3 - 57.5 \%  \\ \cline{1-4}
  LSI      & 82.8 - 76.4 \% & 81.3 - 70.9 \%  & 93.1 - 57.5 \%  \\ \cline{1-4}
 $k$-means & 75.5 - 71.2 \% & 72.8 - 70.8 \%  & 95.3 - 36.2 \%  \\ \cline{1-4}
\end{tabular}
\end{center}
\end{table}

In the second experiment, we evaluate how relevant individual groups of features are to correctly determine cluster equivalence. For this purpose, we tested the accuracies obtained using individual groups of features, as well as using different combination of groups. Since hub CCA had the best  performance of the three algorithms, we used it to compute the values of the cross-lingual article linking features. The results of the evaluation are shown in Table~\ref{table:linkingEval}. We can see that using a single group of features, the highest prediction accuracy can be achieved using  concept related features. The classification accuracy in this case is 88.8\%. By additionally including also the cross-lingual article linking features, the classification accuracy rises slightly to 89.2\%. Using all three groups of features, the achieved accuracy is 89.6\%.

Based on the results, we can make the following conclusions. The cross-lingual similarity algorithms provide valuable information that can be used to identify clusters that describe the same event in different languages. The computed features are however significantly less informative compared to the features computed on the annotated concepts. Nevertheless, the cross-lingual article similarity features are very important for two reasons. The first  is that they allow us to identify for a given cluster a limited set of candidate clusters that are potentially equivalent. This is a very important feature since it reduces the search space by several orders of magnitude. The second reason these features are important is that concept annotations are not available for all articles as the annotation of news articles is computationally intensive and can only be done for a subset of collected articles.

%\begin{table}[h]
%\caption{Accuracy of cluster linking with features obtained from different cross-lingual similarity algorithms}
%\label{table:linkingEvalAlgos}
%\begin{center}
%\begin{tabular}{|c|c|c|c|}
%  \hline
%  \cline{1-4}
%  Models & CA & Precision & Recall \\ \cline{1-4}
%  CCA & 77.6\% & 78.3 \%  & 81.1\%  \\ \cline{1-4}
%  $k$-means & 73.9\% & 69.5\%  & 84.6\%  \\ \cline{1-4}
%  LSI & 79.2\% & 78.9\%  & 80.6\%  \\ \cline{1-4}
%  \hline
%\end{tabular}
%\end{center}
%\end{table}

\begin{table}[h]
\caption{Accuracy of story linking with different sets of features}
\label{table:linkingEval}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Features & CA & Precision & Recall \\ \cline{1-4}
  CCA & 78.2\% & 0.763 & 0.816 \\ \cline{1-4}
  Concepts & 88.8\% & 0.884 & 0.891 \\ \cline{1-4}
  Misc & 65.3\% & 0.718 & 0.500 \\ \cline{1-4}
  CCA + concepts & 89.2\% & 0.891 & 0.893 \\ \cline{1-4}
  All & 89.6\% & 0.895 & 0.895 \\ \cline{1-4}
  \hline
\end{tabular}
\end{center}
\end{table}

\subsection{Scaling}
%\emph{I think that this section should be short.}
%100 languages!
%Memory 100 GB per day
%Newsfeed volume 430k articles per day, 210k english, 36k Spanish, 27k German, 17k French, ....
%Number of similarities, 430k similarities every 200ms, currently could handle 1M articles,
%(n \over 2) 500X500 proxy matrices, 300000x500 Projector matrices, 430000x500 data
%Premultiply with proxy matrices, compare blockwise!
One of the main advantages of our approach is that it is highly scalable. It is fast, very robust to quality of training data, easily extendable, simple to implement and has relatively small hardware requirements. The similarity pipeline is the most computationally intensive part and currently runs on two Intel Xeon E5-2667 v2, 3.30GHz processor machine with 256GB of RAM. This is sufficient to do similarity computation over 150 languages if needed, see \url{http://xling.ijs.si} for live demo over 100 languages. It currently uses Wikipedia as a freely available knowledge base and experiments show that the similarity pipeline dramatically reduces the search space when linking clusters.

Currently we compute similarities over $24$ languages with tags: eng, spa, deu, zho, ita, fra, rus, swe, nld, tur, jpn, por, ara, fin, ron, kor, hrv, tam, hun, slv, pol, srp, cat, ukr but we support any language from the top $100$ Wikipedia languages. Our data stream is Newsfeed (http://newsfeed.ijs.si/) which provides $430k$ unique articles per day. Our system currently computes 2 million similarities per second, that means that we compute $16 \cdot 10^{10}$ similarities per day. We
store one day buffer for each language which requires 1.5 GB of memory with documents   stored as 500 dimensional vectors. We  note that the time complexity of the similarity computations scales linearly with dimension of the feature space and does not  depend on number of languages. For each article, we compute the top $10$  most similar ones in every other language.

For all linear algebra matrix and vector operations, we use high performance numerical linear algebra libraries as BLAS, OPENBLAS and Intel MKL, which currently allows us to process more than one million articles per day.
In our current implementation, we use the variation of hub approach. Our projector matrices are of size $500\times 300,000,$ so every projector takes about $1.1$ GB of RAM. Moreover, we need proxy matrices of size $500\times500$ for every language pair. That is 0.5 GB for $24$ languages and $9.2$ GB for $100$ languages. All together we need around 135 GB of RAM for the system with 100 languages.
 Usage of proxy matrices enables projection of all input documents in the common space and handling language pairs with missing or low alignment. That enables us to do block-wise similarity computations further improving system efficiency. Our code can therefore be easily parallelized using matrix multiplication rather than performing more matrix - vector multiplications. This speeds up our code by a factor around $4.$ In this way, we obtain some caching gains and ability to use vectorization.
Our system is also easily extendable. Adding a new language requires the  computation of  a projector matrix and proxy matrices with all other already available languages.

\section{Discussion and Future Work}
In this paper we have presented a cross-lingual system for linking events in different languages. Building on an existing system, Event Registry, we present and evaluate several approaches to computing a  cross-lingual similarity function. We also present an approach to linking events and evaluate effectiveness of various features.  The  final pipeline is scalable both in terms of number of articles and number of languages, while accurately linking events.

Finally, we discuss some aspects of the methods and the system as a whole.
\begin{itemize}
\item \textbf{LSI and hub CCA} - We first comment on the empirical validation of LSI and hub CCA. On the task of mate retrieval we observe that refining the LSI based projections with hub CCA leads to improved retrieval precision, but the methods perform comparably on the task of event linking. Further inspection showed that the CCA based approach reached a higher precision on smaller clusters. The interpretation is that the linking features are highly aggregated for large clusters, which compensates the lower per-document precision of LSI. Another possible reason is that the advantage that we show on Wikipedia is lost on the news domain - which suggest that domain adaptation should be explored.

\item \textbf{Features} - The experiments show that the hub CCA based features present good baseline, which can greatly benefit from additional semantic extraction based features. Even though the CCA based features do not greatly improve the system performance when the semantic features are used, there are two main benefits in the approach: the linking process can be sped up (smaller candidate cluster sets), and usage in the case when semantic extraction is not possible.

\item \textbf{Future work} - Currently the system is loosely coupled - the language component is built independently of the rest of the system, in particular the linking component. It is possible that better embeddings can be obtained by methods that jointly optimize a classification task and the embedding.

    Another point of interest is to evaluate the system on languages with scarce linguistic resources, where semantic annotation might not be available. For this reason, the labelled dataset of linked clusters should to be extended first. The mate retrieval evaluation showed that even for language pairs with no training set overlap, the hub CCA recovers some signal.

\end{itemize}

%\subsection{Future Work}

%\acks{The authors wish to thank blah blah blah.}

\vskip 0.2in
\bibliography{jair}
\bibliographystyle{theapa}

\end{document}